---
title: "Supervised Feature Selection"
format: html
editor: visual
---

**Load Packages**

```{r}

# Project-relative paths & file management
library(here) 

library(tidyverse)
library(broom)
library(caret)
library(fishualize) 
library(reshape2) # converting correlation matrix to tidy table
library(patchwork) # combine multiple plots
library(infotheo) # calculate mutual information
library(car) # statistics

# wrapper methods
library(randomForest)
library(Boruta)

```

**Load Data as a tibble**

```{r}

stroke_triage_data <- read_csv(here("Code",
                                    "Data Preprocessing",
                                    "data_manipulation_4_UFS",
                                    "data_manipulation_4_UFS.csv"))

```

**Define readable Variable Labels**

```{r}

# Define readable labels
variable_labels <- c(
    gender = "Gender",
    age = "Age",
    bp_systolic = "Systolic BP",
    bp_diastolic = "Diastolic BP",
    pulse = "Pulse",
    blood_glucose = "Blood Glucose",
    temperature = "Temperature",
    vigilance_4ISS = "4I-SS Vigilance",
    gaze_head_deviation_4ISS = "4I-SS Gaze Head Deviation",
    hemiparesis_4ISS = "4I-SS Hemiparesis",
    aphasia_dysarthria_4ISS = "4I-SS Aphasia Dysarthria",
    score_4ISS = "4I-SS Total",
    GCS_eyes = "GCS Eyes",
    GCS_motor = "GCS Motor",
    GCS_verbal = "GCS Verbal",
    score_GCS = "GCS Total",
    NIHSS_at_admission = "NIHSS at Admission",
    medication_anticoags = "Medication Anticoags",
    medication_antiplatelets = "Medication Antiplatelets",
    TOO = "Time of Onset (TOO)",
    exclusion_TOO = "Exclusion TOO",
    exclusion_anticoag_status = "Exclusion Anticoag Status",
    exclusion_surgery_status = "Exclusion Surgery Status",
    DTN = "Door to Needle",
    DTCT = "Door to CT",
    diagnosis_discharge = "Diagnosis at Discharge",
    pre_existing_neurological_deficit = "Pre-existing Neurological Deficit",
    IVT = "IVT",
    EVT = "EVT",
    pupil_reaction_abnormal = "Abnormal Pupil Reaction",
    consciousness_impaired = "Consciousness Impaired",
    mortality = "Mortality",
    infection = "Infection"
  )

```

**Define a Color Palette**

```{r}

# Define the color palette with explicit colors for each category
color_palette <- c(
  "#BC3C29E6",
  "#0072B5E6",
  "#E18727E6",
  "#20854EE6",
  "#7876B1E6",
  "#6F99ADE6",
  "#FFDC91E6",
  "#EE4C97E6"
)

```

# Supervised Feature Selection

**Show variables**

```{r}

str(stroke_triage_data)

```

```{r}

# Transform numeric data into long format, excluding binary and categorical variables
numeric_data <- c("age", "bp_systolic", "pulse", "blood_glucose", "temperature", "TOO")

# Adjust selection of binary variables
binary_vars <- c("gender", "medication_anticoags", "medication_antiplatelets", "exclusion_TOO", "pupil_reaction_abnormal", "pre_existing_neurological_deficit")

# Transform binary and categorical data into long format
binary_categorical_data <- c("vigilance_4ISS", "gaze_head_deviation_4ISS", "hemiparesis_4ISS", "aphasia_dysarthria_4ISS")

# Transform numeric data into long format
melted_numeric_data <- melt(stroke_triage_data[, numeric_data])

# Transform binary and categorical data into long format
melted_binary_categorical_data <- melt(stroke_triage_data[, binary_vars])

# Create density plots for each numeric variable
numeric_plot <- ggplot(melted_numeric_data, aes(x = value)) +
  geom_density(fill = color_palette[3], alpha = 0.7) +
  facet_wrap(~ variable, scales = "free", labeller = as_labeller(variable_labels)) +
  theme_minimal() +
  labs(title = "Distributions of Numeric Variables", x = "Value", y = "Density")

# Create bar plots for each binary and categorical variable
binary_categorical_plot <- ggplot(melted_binary_categorical_data, aes(x = factor(value), fill = factor(value))) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free_x", labeller = as_labeller(variable_labels)) +
  theme_minimal() +
  scale_fill_manual(values = color_palette) +
  labs(title = "Distributions of Binary and Categorical Variables", x = "Value", y = "Count")

# Display plots
print(numeric_plot)
print(binary_categorical_plot)

```

**Set Column Classes**

To ensure that functions work properly, we will first set the column classes for the data. This will ensure that the data is read in as the correct type.

```{r}

stroke_triage_data <- stroke_triage_data %>%  mutate(
    gender = as.factor(gender),
    vigilance_4ISS = as.factor(vigilance_4ISS),
    gaze_head_deviation_4ISS = as.factor(gaze_head_deviation_4ISS),
    hemiparesis_4ISS = as.factor(hemiparesis_4ISS),
    aphasia_dysarthria_4ISS = as.factor(aphasia_dysarthria_4ISS),
    medication_anticoags = as.factor(medication_anticoags),
    medication_antiplatelets = as.factor(medication_antiplatelets),
    exclusion_TOO = as.factor(exclusion_TOO),
    pupil_reaction_abnormal = as.factor(pupil_reaction_abnormal),
    pre_existing_neurological_deficit = as.factor(pre_existing_neurological_deficit),
  )

```

## Filter Methods

### Numerical

**IVT Density Plot**

```{r}

# Convert the data into a long format
long_data <- stroke_triage_data %>%
  select(age, bp_systolic, pulse, blood_glucose, temperature, TOO, IVT) %>%
  pivot_longer(cols = -IVT, names_to = "Variable", values_to = "Value")

# Convert EVT to a factor
long_data$IVT <- as.factor(long_data$IVT)

# Plot with different colors for lines and custom labels for facets
ggplot(long_data, aes(x = Value, color = IVT)) +
  geom_density(size = 1) +
  facet_wrap(~ Variable, scales = "free", labeller = labeller(Variable = variable_labels)) +
  scale_color_manual(values = c("0" = "black", "1" = "darkorange"), labels = c("No", "Yes")) +
  labs(title = "",
       x = "Value",
       y = "Density",
       color = "IVT Therapy") +
  theme_minimal() +
  theme(axis.title = element_text(size = 16),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16),
        axis.text = element_text(size = 14),  
        strip.text = element_text(size = 14),  # Size of facet text
        legend.position = "top") 

```

**IVT Boxplot**

```{r}

# Convert the data into a long format
long_data <- stroke_triage_data %>%
  select(age, bp_systolic, pulse, blood_glucose, temperature, TOO, IVT) %>%
  pivot_longer(cols = -IVT, names_to = "Variable", values_to = "Value")

# Convert IVT to a factor
long_data$IVT <- as.factor(long_data$IVT)

# Plot with different colors for lines and custom labels for facets
ggplot(long_data, aes(x = IVT, y = Value, color = IVT)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free", labeller = labeller(Variable = variable_labels)) +
  scale_color_manual(values = c("0" = "black", "1" = "darkorange"), labels = c("No", "Yes")) +
  labs(x = NULL,
       y = NULL,
       color = "IVT Therapy") +
  theme_minimal() +
  theme(axis.title = element_text(size = 16),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16),
        axis.text = element_text(size = 14),  
        strip.text = element_text(size = 14),  # Size of facet text
        legend.position = "top") 



```

**EVT Density**

```{r}

# Transform data into long format
long_data <- stroke_triage_data %>%
  select(age, bp_systolic, pulse, blood_glucose, temperature, TOO, EVT) %>%
  pivot_longer(cols = -EVT, names_to = "Variable", values_to = "Value")

# Convert EVT to a factor
long_data$EVT <- as.factor(long_data$EVT)

# Create plot with different colors for the lines and custom labels for facets
ggplot(long_data, aes(x = Value, color = EVT)) +
  geom_density(size = 1) +
  facet_wrap(~ Variable, scales = "free", labeller = labeller(Variable = variable_labels)) +
  scale_color_manual(values = c("0" = "black", "1" = "darkred"), labels = c("No", "Yes")) +
  labs(title = "",
       x = "Value",
       y = "Density",
       color = "EVT Therapy") +
  theme_minimal() +
  theme(axis.title = element_text(size = 16),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16),
        axis.text = element_text(size = 14),  
        strip.text = element_text(size = 14),  # Size of facet text
        legend.position = "top")

```

**EVT Boxplot**

```{r}

# Transform data into long format
long_data <- stroke_triage_data %>%
  select(age, bp_systolic, pulse, blood_glucose, temperature, TOO, EVT) %>%
  pivot_longer(cols = -EVT, names_to = "Variable", values_to = "Value")

# Convert EVT to a factor
long_data$EVT <- as.factor(long_data$EVT)

# Create plot with different colors for the lines and custom labels for facets
ggplot(long_data, aes(x = EVT, y = Value, color = EVT)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free", labeller = labeller(Variable = variable_labels)) +
  scale_color_manual(values = c("0" = "black", "1" = "darkred"), labels = c("No", "Yes")) +
  labs(x = NULL,
       y = NULL,
       color = "EVT Therapy") +
  theme_minimal() +
  theme(axis.title = element_text(size = 16),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 16),
        axis.text = element_text(size = 14),  
        strip.text = element_text(size = 14),  # Size of facet text
        legend.position = "top") 


```

**Normal Distribution Test**

To determine which test to use, we will first check the normality assumptions. Based on the results, we will decide whether to use a parametric test or not.

```{r}

# choose numerical variables

numeric_variables <- stroke_triage_data %>%
  select(age, bp_systolic, pulse, blood_glucose, temperature, TOO, EVT)

```

```{r}

# Perform Shapiro-Wilk tests and collect results
shapiro_results <- numeric_variables %>%
  select(-EVT) %>%
  map_df(~ broom::tidy(shapiro.test(.)), .id = "Variable")

# Format the results
shapiro_results <- shapiro_results %>%
  mutate(
    p_value_formatted = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ sub("^0\\.", ".", format(round(p.value, 3), nsmall = 3))
    ),
    statistic = round(statistic, 3)
  ) %>%
  select(Variable, statistic, p_value = p_value_formatted) %>%
  arrange(p_value)

# Display results
print(shapiro_results)

```

**Q-Q Plots**

```{r}

qq_test_variables <- numeric_variables %>%
  select(-EVT) 

# Melt the data into a long format
long_data <- qq_test_variables %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value")

# Create Q-Q plot with facets
qq_plot <- ggplot(long_data, aes(sample = Value)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~ Variable, scales = "free", labeller = as_labeller(variable_labels)) +
  ggtitle("") +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

# Display the plot
print(qq_plot)

```

**Mann-Whitney U-Test**

```{r}

# Perform Mann-Whitney U tests and calculate medians
mann_whitney_results <- numeric_variables %>%
  select(-EVT) %>%
  map2_dfr(names(.), ~{
    test <- wilcox.test(.x ~ numeric_variables$EVT)
    data.frame(
      Variable = .y,
      median_no_EVT = median(.x[numeric_variables$EVT == 0], na.rm = TRUE),
      median_EVT = median(.x[numeric_variables$EVT == 1], na.rm = TRUE),
      median_diff = median(.x[numeric_variables$EVT == 1], na.rm = TRUE) - median(.x[numeric_variables$EVT == 0], na.rm = TRUE),
      u_statistic = test$statistic,
      p_value = test$p.value
    )
  })

# Add readable variable labels and format p-values
mann_whitney_results <- mann_whitney_results %>%
  mutate(
    Variable = variable_labels[Variable],
    p_value_formatted = case_when(
      p_value < 0.001 ~ "< .001",
      TRUE ~ sub("^0\\.", ".", sprintf("%.3f", p_value))
    )
  )

# Select and rename columns, then sort
mann_whitney_results <- mann_whitney_results %>%
  select(Variable, median_no_EVT, median_EVT, median_diff, u_value = u_statistic, p_value = p_value_formatted) %>%
  arrange(p_value)

# Round the numeric columns to two decimal places
mann_whitney_results <- mann_whitney_results %>%
  mutate(across(c(median_no_EVT, median_EVT, median_diff, u_value), ~ round(., 2)))

# Print the results
print(mann_whitney_results)

```

### Binary

```{r}

stroke_triage_data$EVT <- as_factor(stroke_triage_data$EVT)

```

**Chi-Squared Test Binary Variables**

```{r}

# List of binary variables
binary_variables <- c("gender", "medication_anticoags", "medication_antiplatelets", 
                      "exclusion_TOO", "pupil_reaction_abnormal", 
                      "pre_existing_neurological_deficit")
 

# Perform Chi-Square tests and combine results
chi_square_results <- map_dfr(binary_variables, function(var) {
  test <- chisq.test(table(stroke_triage_data[[var]], stroke_triage_data$EVT))
  tidy_test <- tidy(test)
  
  # Add readable variable name
  tidy_test <- tidy_test %>%
    mutate(Variable = variable_labels[var])
  
  return(tidy_test)
})

# Format results according to APA guidelines
chi_square_results <- chi_square_results %>%
  select(Variable, chi_statistic = statistic, p.value) %>%
  mutate(
    p_value_formatted = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ sub("^0\\.", ".", format(round(p.value, 3), nsmall = 3))
    ),
    chi_statistic = round(chi_statistic, 2)
  ) %>%
  select(Variable, chi_statistic, p_value = p_value_formatted) %>%
  arrange(p_value)

# Display results
print(chi_square_results)

```

**Mutual Information (Binary)**

```{r}

# Perform MI tests and combine results
mi_results <- map_dfr(binary_variables, function(var) {
  # Convert variable to factor if not already
  stroke_triage_data[[var]] <- as.factor(stroke_triage_data[[var]])
  
  # Calculate MI
  mi <- mutinformation(stroke_triage_data[[var]], stroke_triage_data$EVT)
  
  # Create a tidy dataframe for the results
  tidy_mi <- tibble(
    Variable = variable_labels[var],
    mi_value = mi 
  )
  
  return(tidy_mi)
})

# Format MI values without scientific notation and sort in descending order
mi_results <- mi_results %>%
  mutate(
    mi_value = format(as.numeric(mi_value), scientific = FALSE, digits = 5)
  ) %>%
  arrange(desc(as.numeric(mi_value)))

# Display results
print(mi_results)

```

### Ordinal

```{r}

# Manually reorder levels in the dataset
visualize_data <- stroke_triage_data %>%
  mutate(
    vigilance_4ISS = factor(vigilance_4ISS, levels = c("2", "1", "0")),
    gaze_head_deviation_4ISS = factor(gaze_head_deviation_4ISS, levels = c("2", "1", "0")),
    hemiparesis_4ISS = factor(hemiparesis_4ISS, levels = c("2", "1", "0")),
    aphasia_dysarthria_4ISS = factor(aphasia_dysarthria_4ISS, levels = c("1", "0")),
    EVT = factor(EVT, levels = c("1", "0"), labels = c("EVT", "No EVT"))
  )

```

```{r}

gray_palette <- c("2" = "#4D4D4D", "1" = "#B3B3B3", "0" = "#E0E0E0")

# Define readable labels
variable_labels <- c(
  vigilance_4ISS = "4I-SS Vigilance",
  gaze_head_deviation_4ISS = "4I-SS Gaze Head Deviation",
  hemiparesis_4ISS = "4I-SS Hemiparesis",
  aphasia_dysarthria_4ISS = "4I-SS Aphasia Dysarthria"
)

# Create a function for horizontal stacked bar plots with gray scales and no legend
plot_horizontal_stacked_bar <- function(data, variable, label) {
  ggplot(data, aes_string(x = label, fill = variable)) +
    geom_bar(position = "fill", color = "black", size = 1.0) +
    scale_fill_manual(values = gray_palette) +
    labs(title = variable_labels[[variable]], x = NULL, y = NULL, fill = NULL) +
    scale_y_continuous(labels = scales::percent) +
    coord_flip() +
    theme_minimal() +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          plot.title = element_text(hjust = 0.5, size = 11),
          legend.position = "none")
}

# Plot for each 4I-SS variable
p1 <- plot_horizontal_stacked_bar(visualize_data, "vigilance_4ISS", "EVT")
p2 <- plot_horizontal_stacked_bar(visualize_data, "gaze_head_deviation_4ISS", "EVT")
p3 <- plot_horizontal_stacked_bar(visualize_data, "hemiparesis_4ISS", "EVT")
p4 <- plot_horizontal_stacked_bar(visualize_data, "aphasia_dysarthria_4ISS", "EVT")

# Combine plots using patchwork, with no legends
combined_plot <- (p1 + p2) / (p3 + p4)

# Display the combined plot
print(combined_plot)

```

**Chi-Squared Test 4I-SS Elements**

```{r}

# List of variables
stroke_scale_variables <- c("vigilance_4ISS",
                            "gaze_head_deviation_4ISS",
                            "hemiparesis_4ISS",
                            "aphasia_dysarthria_4ISS")


# Perform chi-square tests and combine results
chi_square_results <- map_dfr(stroke_scale_variables, function(var) {
  test <- chisq.test(table(stroke_triage_data[[var]], stroke_triage_data$EVT))
  tidy_test <- tidy(test)
  
  # Add readable variable name
  tidy_test <- tidy_test %>%
    mutate(Variable = variable_labels[var])
  
  return(tidy_test)
})

# Format results according to APA guidelines
chi_square_results <- chi_square_results %>%
  select(Variable, chi_statistic = statistic, p.value) %>%
  mutate(
    p_value_formatted = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ sub("^0\\.", ".", format(round(p.value, 3), nsmall = 3))
    ),
    chi_statistic = round(chi_statistic, 2)
  ) %>%
  select(Variable, chi_statistic, p_value = p_value_formatted) %>%
  arrange(p_value)

# Display results
print(chi_square_results)
  
```

**Mutual Information (4I-SS Elements)**

```{r}

# Perform MI tests and combine results
mi_results <- map_dfr(stroke_scale_variables, function(var) {
  # Convert variable to factor if not already
  stroke_triage_data[[var]] <- as.factor(stroke_triage_data[[var]])
  
  # Calculate MI
  mi <- mutinformation(stroke_triage_data[[var]], stroke_triage_data$EVT)
  
  # Create a tidy dataframe for the results
  tidy_mi <- tibble(
    Variable = variable_labels[var],
    mi_value = mi  # Do not round here
  )
  
  return(tidy_mi)
})

# Format MI values without scientific notation and sort in descending order
mi_results <- mi_results %>%
  mutate(
    mi_value = format(as.numeric(mi_value), scientific = FALSE, digits = 5)
  ) %>%
  arrange(desc(as.numeric(mi_value)))

# Display results
print(mi_results)

```

### Logistic Regression

```{r}

# Train the logistic regression model
model <- glm(EVT ~ age + bp_systolic + pulse + blood_glucose + temperature + TOO + 
             gender + medication_anticoags + medication_antiplatelets + 
             exclusion_TOO + pupil_reaction_abnormal + pre_existing_neurological_deficit +
             vigilance_4ISS + gaze_head_deviation_4ISS + hemiparesis_4ISS + aphasia_dysarthria_4ISS, 
             data = stroke_triage_data, 
             family = binomial)

# Extract the model summary and coefficients matrix
summary_model <- summary(model)
coefficients <- summary_model$coefficients

# Remove the intercept row by using its row name
coefficients <- coefficients[rownames(coefficients) != "(Intercept)", ]

# Calculate Wald statistics
wald_stats <- (coefficients[, "Estimate"] / coefficients[, "Std. Error"])^2

# Calculate odds ratios
odds_ratios <- exp(coefficients[, "Estimate"])

# Calculate 95% confidence intervals and remove the intercept using row names
conf_int <- exp(confint(model))
conf_int <- conf_int[rownames(coefficients), ]

# (Optional) Clean up predictor names if needed (e.g., remove trailing digits)
rownames(coefficients) <- gsub("1$", "", rownames(coefficients))

# Combine all information into a single data frame
results <- data.frame(
  B = coefficients[, "Estimate"],
  SE = coefficients[, "Std. Error"],
  Wald = wald_stats,
  p = coefficients[, "Pr(>|z|)"],
  `Exp(B)` = odds_ratios,
  `95% CI Lower` = conf_int[, 1],
  `95% CI Upper` = conf_int[, 2]
)

# Display the results
print(results)

```

**Variance Inflation Factors**

Before performing the logistic regression, the Variance Inflation Factor (VIF) was evaluated for each predictor to assess multicollinearity. The VIF values for the predictors ranged from 1.08 to 2.25, which are below the advised threshold of 10. This indicates that multicollinearity is not a significant issue in this model, allowing for reliable estimates of the coefficients.

```{r}

# Fit a reduced model excluding pupil_reaction_abnormal
model_vif <- glm(EVT ~ age + bp_systolic + pulse + blood_glucose + temperature + 
                 TOO + gender + medication_anticoags + medication_antiplatelets + 
                 exclusion_TOO + pre_existing_neurological_deficit +
                 vigilance_4ISS + gaze_head_deviation_4ISS + hemiparesis_4ISS + 
                 aphasia_dysarthria_4ISS, 
                 data = stroke_triage_data, 
                 family = binomial)

# Calculate VIF values (only for multicollinearity check)
vif_values <- vif(model_vif)

# Show results
print(vif_values)

```

### Stacked Bar Chart of Feature Relevance Assessment

#### Numerical Variables

**Calculate the Boundaries of the Groups**

Use the min max values to separate the scale into x groups.

```{r}

stroke_triage_data %>%
  select(age, bp_systolic, pulse, blood_glucose, temperature, TOO) %>%
  summarise(across(everything(), list(min = ~min(.x, na.rm = TRUE), max = ~max(.x, na.rm = TRUE))))

```

```{r}

# Given min and max values for each variable
age_min <- 20
age_max <- 99
bp_systolic_min <- 84
bp_systolic_max <- 250
pulse_min <- 42
pulse_max <- 198
blood_glucose_min <- 40
blood_glucose_max <- 485
temperature_min <- 34.4
temperature_max <- 39.1
TOO_min <- 0.2
TOO_max <- 46.8

# Number of groups for each variable
n_age_groups <- 5
n_bp_systolic_groups <- 4
n_pulse_groups <- 4
n_blood_glucose_groups <- 4
n_temperature_groups <- 2
n_TOO_groups <- 4

# Define the intervals for each category, rounding to the nearest whole number
age_breaks <- round(seq(age_min, age_max, length.out = n_age_groups + 1))
bp_systolic_breaks <- round(seq(bp_systolic_min, bp_systolic_max, length.out = n_bp_systolic_groups + 1))
pulse_breaks <- round(seq(pulse_min, pulse_max, length.out = n_pulse_groups + 1))
blood_glucose_breaks <- round(seq(blood_glucose_min, blood_glucose_max, length.out = n_blood_glucose_groups + 1))
temperature_breaks <- round(seq(temperature_min, temperature_max, length.out = n_temperature_groups + 1), 1)
TOO_breaks <- round(seq(TOO_min, TOO_max, length.out = n_TOO_groups + 1), 1)

# Create labels based on these intervals
age_labels <- paste(head(age_breaks, -1), "-", tail(age_breaks, -1), sep = "")
bp_systolic_labels <- paste(head(bp_systolic_breaks, -1), "-", tail(bp_systolic_breaks, -1), sep = "")
pulse_labels <- paste(head(pulse_breaks, -1), "-", tail(pulse_breaks, -1), sep = "")
blood_glucose_labels <- paste(head(blood_glucose_breaks, -1), "-", tail(blood_glucose_breaks, -1), sep = "")
temperature_labels <- paste(head(temperature_breaks, -1), "-", tail(temperature_breaks, -1), sep = "")
TOO_labels <- paste(head(TOO_breaks, -1), "-", tail(TOO_breaks, -1), sep = "")

# Create variable labels with units for numeric variables
numeric_variable_labels <- c(
    age_groups = "Age",
    bp_systolic_groups = "Systolic BP (mmHg)",
    pulse_groups = "Heart Rate",
    blood_glucose_groups = "Blood Glucose (mg/dL)",
    temperature_groups = "Body Temperature (°C)",
    TOO_groups = "Time of Onset (Hours)"
)

# Create a new DataFrame and apply categorizations using the custom breaks
stroke_triage_data_grouped <- stroke_triage_data %>%
  mutate(age_groups = factor(cut(age, breaks = age_breaks, labels = age_labels, include.lowest = TRUE)),
         bp_systolic_groups = factor(cut(bp_systolic, breaks = bp_systolic_breaks, labels = bp_systolic_labels, include.lowest = TRUE)),
         pulse_groups = factor(cut(pulse, breaks = pulse_breaks, labels = pulse_labels, include.lowest = TRUE)),
         blood_glucose_groups = factor(cut(blood_glucose, breaks = blood_glucose_breaks, labels = blood_glucose_labels, include.lowest = TRUE)),
         temperature_groups = factor(cut(temperature, breaks = temperature_breaks, labels = temperature_labels, include.lowest = TRUE)),
         TOO_groups = factor(cut(TOO, breaks = TOO_breaks, labels = TOO_labels, include.lowest = TRUE))) 

# Convert the data into long format for numeric variables
stroke_triage_data_numeric_long <- stroke_triage_data_grouped %>%
  pivot_longer(cols = c(age_groups, bp_systolic_groups, pulse_groups, blood_glucose_groups, temperature_groups, TOO_groups),
               names_to = "name",
               values_to = "value")

# Convert EVT to a factor with labels "No" and "Yes"
stroke_triage_data_numeric_long$EVT <- factor(stroke_triage_data_numeric_long$EVT, levels = c(0, 1), labels = c("No", "Yes"))


ggplot(stroke_triage_data_numeric_long, aes(y = value, fill = EVT)) +
  geom_bar(position = "fill") +
  facet_wrap(vars(name), scales = "free", ncol = 2, labeller = labeller(name = numeric_variable_labels)) +
  scale_fill_manual(values = c("Yes" = "darkgrey", "No" = "white")) +
  theme(
    strip.text = element_text(size = 14), 
    legend.text = element_text(size = 14), 
    legend.title = element_text(size = 14), 
    legend.position = "top",
    axis.text.y = element_text(size = 14) 
  ) +
  labs(x = NULL, y = NULL, fill = "EVT Therapy")


```

**How Many Instances are in Specific Groups**

Some groups may have a limited number of instances, this could potentially lead to misinterpretation.

```{r}

# List of groups to analyze
groups <- c("age_groups", "bp_systolic_groups", "pulse_groups", "blood_glucose_groups", "temperature_groups", "TOO_groups")

# Create an empty list to store the results
group_counts <- list()

# Loop through each group and count occurrences
for (group in groups) {
  group_counts[[group]] <- stroke_triage_data_grouped %>%
    group_by(!!sym(group)) %>%
    summarise(count = n()) %>%
    arrange(desc(count))
}

# Display the counts for each group
group_counts

```

#### Binary Variables

```{r}

# categorical and binary variables
categorical_binary_variable_labels <- c(
    exclusion_TOO = "Clinical Exclusion TOO",
    medication_anticoags = "Medication Anticoags",
    medication_antiplatelets = "Medication Antiplatelets",
    pupil_reaction_abnormal = "Abnormal Pupil Reaction",
    pre_existing_neurological_deficit = "Pre-existing Neurological Deficit",
    gender = "Gender"
)

# Filter variables to include in the plot
stroke_triage_data_categorical_long <- stroke_triage_data_grouped %>%
  pivot_longer(cols = c(exclusion_TOO, medication_anticoags, medication_antiplatelets, pupil_reaction_abnormal, pre_existing_neurological_deficit, gender),
               names_to = "name",
               values_to = "value")

# Convert EVT to a factor with labels "No" and "Yes"
stroke_triage_data_categorical_long$EVT <- factor(stroke_triage_data_categorical_long$EVT, levels = c(0, 1), labels = c("No", "Yes"))

# Plot for the selected categorical and binary variables
ggplot(stroke_triage_data_categorical_long, aes(y = value, fill = EVT)) +
  geom_bar(position = "fill") +
  facet_wrap(vars(name), scales = "free", ncol = 2, labeller = labeller(name = categorical_binary_variable_labels)) +
  scale_fill_manual(values = c("Yes" = "darkgrey", "No" = "white")) + 
  theme(
    strip.text = element_text(size = 14),
    legend.text = element_text(size = 14), 
    legend.title = element_text(size = 14), 
    legend.position = "top" 
  ) +
  labs(x = NULL, y = NULL, fill = "EVT Therapy")

```

#### 4I-SS Assessments

```{r}

# Define variable labels for 4I-SS items
four_iss_variable_labels <- c(
    vigilance_4ISS = "4I-SS Vigilance",
    gaze_head_deviation_4ISS = "4I-SS Gaze Head Deviation",
    hemiparesis_4ISS = "4I-SS Hemiparesis",
    aphasia_dysarthria_4ISS = "4I-SS Aphasia Dysarthria"
)

# Convert the data into long format for 4I-SS items
stroke_triage_data_4iss_long <- stroke_triage_data_grouped %>%
  pivot_longer(cols = c(vigilance_4ISS, gaze_head_deviation_4ISS, hemiparesis_4ISS, aphasia_dysarthria_4ISS),
               names_to = "name",
               values_to = "value")

# Convert EVT to a factor with labels "No" and "Yes"
stroke_triage_data_4iss_long$EVT <- factor(stroke_triage_data_4iss_long$EVT, levels = c(0, 1), labels = c("No", "Yes"))

# Plot for 4I-SS items
ggplot(stroke_triage_data_4iss_long, aes(y = value, fill = EVT)) +
  geom_bar(position = "fill") +
  facet_wrap(vars(name), scales = "free", ncol = 2, labeller = labeller(name = four_iss_variable_labels)) +
  scale_fill_manual(values = c("Yes" = "darkgrey", "No" = "white")) +
  theme(
    strip.text = element_text(size = 14), 
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14), 
    legend.position = "top" 
  ) +
  labs(x = NULL, y = NULL, fill = "EVT Therapy") 

```

Test, if this plot is plausible:

```{r}

stroke_triage_data %>%
  group_by(gaze_head_deviation_4ISS) %>%
  summarise(
    total = n(),
    evt_count = sum(EVT == 1),
    evt_percentage = (evt_count / total) * 100
  )

```

## Wrapper Methods

The interpretation of accuracy and many other metrics is highly influenced by class imbalances. In our dataset, the positive EVT class constitutes only 6% of the total samples. Most feature selection wrapper methods are based on accuracy to identify the most significant features. In an unbalanced dataset, the feature selection process is biased towards predicting the majority class (No EVT), which contradicts our goal. This can lead to a bias towards predicting the "No EVT" class due to its majority presence. However, the aim is to identify variables that can accurately predict the EVT class.

To preserve real-world conditions for model training and evaluation, the class distribution of the training dataset will not be adjusted. For the purpose of feature selection, the EVT class will be balanced to a 50/50 ratio with a total of 600 observations. This involves a combination of oversampling and undersampling techniques. Pure oversampling would generate an excessive number of instances from the 77 positive EVT data points, while pure undersampling would reduce the dataset to only 144 instances. This balanced dataset will enable us to effectively use accuracy-based feature selection packages and understand the influence of each feature on the EVT label in a multivariate context.

**Sampling Methods:**

```{r}

# Sample 300 positive instances with replacement, because we have less than 300 observations
positive_samples <- stroke_triage_data %>% filter(EVT == 1) %>% sample_n(300, replace = TRUE)

# Sample 300 negative instances without replacement, because we have more than 300 observations
negative_samples <- stroke_triage_data %>% filter(EVT == 0) %>% sample_n(300)

# Combine the positive and negative samples into a balanced DataFrame
stroke_triage_data_balanced <- bind_rows(positive_samples, negative_samples)

# Check the distribution of the classes
table(stroke_triage_data_balanced$EVT)

```

**Three Approaches:**

1.  `caret:` Recursive Feature Elimination (RFE)

2.  `mlr:` Sequential Feature Selection

3.  `Boruta:` for Feature Selection

### `caret`

Recursive Feature Elimination (RFE) is a backward selection technique used for feature selection. In this approach, models are iteratively trained by removing the least important features at each step. The model used in our case was a decision tree model.

RFE does not test every possible combination of features. Instead, it starts with all features and systematically removes the least important ones, evaluating the model's performance with different subsets of features using cross-validation. The result is an evaluation of how different subsets of features affect the model's performance, typically measured by accuracy or another performance metric. Cross-validation is employed to ensure robust and reliable performance estimates.

**Train a Random Forest**

```{r}

# Define the control using a backward selection method with decision trees
control <- caret::rfeControl(
  functions = caret::rfFuncs, # Using random forest functions for RFE
  method = "cv",              # Applying cross-validation
  number = 10,                # Using 10-fold cross-validation
  saveDetails = TRUE          # Saving detailed information for further analysis
)

# Run the RFE algorithm
results <- caret::rfe(
  stroke_triage_data_balanced[, -which(names(stroke_triage_data_balanced) == "EVT")], # Excluding the target variable 'EVT' from the predictors
  stroke_triage_data_balanced$EVT,                                                    # Defining the target variable
  sizes = c(1:16),                                                                    # Evaluating feature subsets ranging from 1 to 16 variables
  rfeControl = control,                                                               # Applying the previously defined control parameters
  method = "rpart"                                                                    # Using decision trees for RFE
)

```

**Printing**

The *caret::rfe* function identifies the most important features. In our case, the most important features were severity of hemiparesis (as measured by the 4-Item Stroke Scale), Time Of Onset, heart rate, patient's age, and systolic blood pressure. The function recommended using 12 out of 16 features, but not information about a pre-existing neurological deficit and information of antiplatelet medication.

```{r}

print(results)

```

**Plot the Results**

```{r}

plot(results, type = c("g", "o"))

```

### `Boruta`

Boruta is a wrapper method that assigns an importance score to each feature in the dataset. Unlike other backward selection techniques, it does not eliminate features but ranks them all. In this thesis, 100 random forest models were trained to compare the mean decrease accuracy of actual features to that of shadow features. The shadow features contain the same values as the actual features but are randomly shuffled. This comparison calculates a z-score, which is plotted on the vertical axis to represent the importance of each feature relative to others. However, this plot only allows for a relative comparison within the features and does not interpret the actual effect. Due to the non-deterministic nature of random forest classifiers and the shuffling process, the results slightly vary with each iteration. The different results are represented using boxplots.

**Create `Boruta` Model**

```{r}

# Set the seed for reproducibility
set.seed(456)

# Apply the Boruta algorithm
boruta_output <- Boruta(EVT ~ .,   # EVT is the label, and . indicates that all other variables in the dataset are predictors
                        data = stroke_triage_data_balanced)   # The dataset used for the Boruta algorithm


```

**How Many Random Forest Models were Created?**

```{r}

# How many Random Forrest were created
print(boruta_output$maxRuns)

```

**Plot Boruta**

```{r}

# Define more readable labels
label_names <- c("Shadow Min",
                 "Shadow Mean",
                 "Shadow Max",
                 "Neurolog. Deficit",
                 "Abnormal Pupils",
                 "Anticoagulants",
                 "Antiplatelets",
                 "Gender",
                 "Speech",
                 "Exclusion TOO",
                 "Gaze Head Deviation",
                 "Vigilance",
                 "BP Systolic",
                 "Temperature",
                 "Pulse",
                 "Age",
                 "Blood Glucose",
                 "Time of Onset",
                 "Hemiparesis")

# Plot the Boruta output without x-axis and y-axis labels and ticks
plot(boruta_output,
     xlab = "",   # Remove x-axis label
     ylab = "",
     xaxt = "n",  # Do not draw x-axis ticks and labels
     yaxt = "n")  # Do not draw y-axis ticks and labels)

# Add the x-axis without labels
axis(1,
     at = 1:length(label_names),  # Position ticks at each label position
     labels = FALSE)  # Do not draw labels on the x-axis

# Add the rotated labels
text(x = 1:length(label_names),  # x-coordinates for the labels
     y = par("usr")[3] - 3,  # y-coordinate for the labels, slightly below the plot
     srt = 45,  # Rotate the text 45 degrees
     adj = 1,  # Right-align the text
     labels = label_names,  # The labels to be displayed
     xpd = TRUE,  # Allow text to be drawn outside the plot region
     cex = 0.7)  # Scale the size of the labels

# Scale down the y-axis labels
axis(2,
     cex.axis = 0.7)  # Scale down the size of the y-axis labels

```

## Intrinsic Methods

We use the intrinsic properties of `randomForest` to evaluate the importance of the features.

```{r}

# Train Random Forest model
rf_model <- randomForest(EVT ~ ., data = stroke_triage_data_balanced, importance = TRUE)

# Calculate feature importance
importance_values <- importance(rf_model)

# Display feature importance
print(importance_values)

# Plot feature importance
varImpPlot(rf_model)

```

**Export Data for Machine Learning**

Export the ML-Ready Data to `data_manipulation_6_SFS`.

```{r}

write_csv(stroke_triage_data, "data_manipulation_6_SFS.csv")

```

**Export Data for Exploratory Data Analysis**

```{r}

write_csv(stroke_triage_data, here("Code", "Exploratory Data Analysis", "stroke_triage_data_preprocessed.csv"))

```
