---
title: "Exploratory Data Analysis"
editor: visual
---

## Load Packages

```{r message = FALSE, warning = FALSE}

library(here) # Project-relative paths & file management

library(DT) # Present Tables using DT

library(tidyverse) # data manipulation
library(ggpubr) # Based Publication Ready Plots
library(crosstable) # Create easy crosstables
library(rsample) # For the train- test split
library(flextable) # To safe the crosstable html output

library(tidymodels) # For Imputation

library(gridExtra) # Display multiple plots
 
```

**Define readable Variable Labels**

```{r}

# Define readable labels
variable_labels <- c(
  gender = "Gender",
  age = "Age",
  bp_systolic = "Systolic BP",
  bp_diastolic = "Diastolic BP",
  pulse = "Pulse",
  blood_sugar = "Blood Glucose",
  temperature = "Temperature",
  vigilance_4ISS = "4I-SS Vigilance",
  gaze_head_deviation_4ISS = "4I-SS Gaze Head Deviation",
  hemiparesis_4ISS = "4I-SS Hemiparesis",
  aphasia_dysarthria_4ISS = "4I-SS Aphasia Dysarthria",
  score_4ISS = "4I-SS Total",
  GCS_eyes = "GCS Eyes",
  GCS_motor = "GCS Motor",
  GCS_verbal = "GCS Verbal",
  score_GCS = "GCS Total",
  NIHSS_at_admission = "NIHSS at Admission",
  medication_anticoags = "Medication Anticoags",
  medication_antiplatelets = "Medication Antiplatelets",
  TOO = "Time of Onset (TOO)",
  exclusion_TOO = "Exclusion TOO",
  exclusion_anticoag_status = "Exclusion Anticoag Status",
  exclusion_surgery_status = "Exclusion Surgery Status",
  DTN = "Door to Needle",
  DTCT = "Door to CT",
  diagnosis_discharge = "Diagnosis at Discharge",
  pre_existing_neurological_deficit = "Pre-existing Neurological Deficit",
  IVT = "IVT",
  EVT = "EVT",
  pupil_reaction_abnormal = "Abnormal Pupil Reaction",
  consciousness_impaired = "Consciousness Impaired",
  mortality = "Mortality",
  infection = "Infection",
  year_of_arrival = "Year",
  diagnosis_discharge = "Diagnosis at Discharge",
  ways_of_admission_transport = "Type of Transport",
  ways_of_admission_origin = "Type of Origin"
)

```

# Unmatched Data

**Load data**

```{r}

# Load unmatched EMS data
ems_data <- read_csv(here("Data", "ems_data_original.csv"))

# Load unmatched clinical data
clinical_data <- read_csv(here("Data", "stroke_angel_data_all.csv"))

```

**Convert missing values to NaN**

```{r}

# List of values to be treated as NA
missing_values <- c("", " ", "NA", "N/A", "n/a", "na", "nan", "NaN", "null", "Null", "NULL", "None", "none", "-", -4, -1)

# Replace missing values in the entire EMS DataFrame
ems_data <- ems_data %>%
  mutate(across(everything(), ~replace(., . %in% missing_values, NA)))

# Replace missing values in the entire Clinical DataFrame
clinical_data <- clinical_data %>%
  mutate(across(everything(), ~replace(., . %in% missing_values, NA)))

```

**Remove Multiple Patient Entries in EMS-Data**

Many patients have not one but multiple entries, so we keep the entry with the most information and drop the others to have a tidy format. See the detailed explanation of this step in the script `data_manipulation_1_merge` in the section 'Remove Multiple Patient Entries in EMS-Data'.

```{r}
#| warning: false

ems_data <- ems_data %>% 
  mutate(ankunftszeit_laut_voranmeldung = ymd_hms(ankunftszeit_laut_voranmeldung))

ems_data$ankunftszeit_rounded <- floor_date(ems_data$ankunftszeit_laut_voranmeldung, "hour")

ems_data <- ems_data %>% select(-"ankunftszeit_laut_voranmeldung") %>% distinct(.keep_all = TRUE)

# Create IDs (rounded arrival-time + age)
ems_data$matching_ID_ems <- paste(ems_data$ankunftszeit_rounded, ems_data$alter, sep = "_")


# write function
prioritize_complete_entries <- function(ems_data) {
  ems_data %>%
    # Calculate the count of non-missing values for each row
    mutate(non_null_count = rowSums(!is.na(select(., -matching_ID_ems)), na.rm = TRUE)) %>%
    # Group by matching ID
    group_by(matching_ID_ems) %>%
    # Sort to prioritize the most complete and most recent entry within each group
    arrange(desc(non_null_count), desc(ankunftszeit_rounded)) %>%
    slice(1) %>%
    ungroup() %>%
    select(-non_null_count)
}

# Apply the function to ems_data
ems_data <- prioritize_complete_entries(ems_data)

```

**Set correct data types**

```{r}

# Prepare EMS Data
ems_data <- ems_data %>%
  mutate(
    # Set to integer
    age = as.integer(parse_number(alter)), 
    pulse = as.integer(puls),
    bp_systolic = as.integer(bd_systolisch),
    bp_diastolic = as.integer(bd_diastolisch),
    blood_glucose = as.integer(bz),
    temperature = as.integer(temperatur),
    
    # Extract first digit for 4ISS score (its a string right now)
    vigilance_4ISS = factor(str_sub(vigilanz_4ISS, 1, 1),
                            levels = c("0", "1", "2")),
    gaze_head_deviation_4ISS = factor(str_sub(kopf_blickwendung_4ISS, 1, 1),
                                      levels = c("0", "1", "2")),
    hemiparesis_4ISS = factor(str_sub(hemiparese_4ISS, 1, 1),
                              levels = c("0", "1", "2")),
    
    # Rename aphasia/dysarthria status
    aphasia_dysarthria_4ISS = factor(case_when(
      str_detect(sprach_sprechstoerung_4ISS, "ja") ~ "1",
      str_detect(sprach_sprechstoerung_4ISS, "nein") ~ "0",
      TRUE ~ NA_character_)
      ))


# Prepare Clinical Data 
clinical_data <- clinical_data %>%
  mutate(
    # Set to factor
    gender = as_factor(sex),
    medication_anticoags = as_factor(ly_med_akoag),
    exclusion_TOO = as_factor(ly_med_akoag),
    IVT = as_factor(ly_lyse),
    EVT = as_factor(ly_thromb),
    
    # Set to integer
    TOO = as.integer(ly_sym_dauer),
    
    # Rename NIHSS variable
    NIHSS_at_admission = ly_nihssbauf,
    
    # Extract admission year from date string
    year_of_arrival = as_factor(year(mdy(a_dat))),
    
    # Recode transport categories
    ways_of_admission_transport = recode_factor(
      as.character(ly_einlief),
      "2"  = "EMS with Emergency Physician",
      "3"  = "EMS without Emergency Physician",
      "13" = "EMS with Emergency Physician",
      "4"  = "EMS with Emergency Physician",
      "9"  = "EMS without Emergency Physician",
      "10" = "Other Transport",
      "11" = "Other Transport",
      "5"  = "EMS without Emergency Physician",
      "6"  = "EMS with Emergency Physician",
      "7"  = "Other Transport",
      "8"  = "Other Transport",
      "12" = "Other Transport",
      "1"  = "Self Instruction"
    ),
    
    # Recode admission origin categories
    ways_of_admission_origin = recode_factor(
      as.character(ly_einlief),
      "2"  = "EMS via Emergency Call",
      "3"  = "EMS via Emergency Call",
      "13" = "EMS via Emergency Call",
      "4"  = "General Practitioner",
      "9"  = "General Practitioner",
      "10" = "General Practitioner",
      "11" = "General Practitioner",
      "5"  = "Drip and Ship",
      "6"  = "Drip and Ship",
      "7"  = "Drip and Ship",
      "8"  = "Drip and Ship",
      "12" = "Internal Transfer",
      "1"  = "Self Instruction"
    ),
    
    # Recode discharge diagnosis 
    diagnosis_discharge = recode_factor(
      as.character(ly_ediag),
      "1" = "Stroke",
      "2" = "TIA",
      "3" = "TGA",
      "4" = "Bleeding",
      "5" = "Traumatic Brain Injury",
      "6" = "SAB",
      "7" = "SDH/Hematoma",
      "8" = "Stroke Mimic"
    ))

```

**Bind Clinical and EMS data into one unmatched Dataframe**

```{r}

unmatched_data <- bind_rows(ems_data, clinical_data) 

```

# Matched Data

**Load data**

```{r}

matched_data <- read_csv(here("Code",
                              "Exploratory Data Analysis",
                              "stroke_triage_data_raw.csv"))

```

**Set correct data types**

```{r}

matched_data <- matched_data %>%
  mutate(
    gender = as.factor(gender),
    age = as.integer(age),
    bp_systolic = as.integer(bp_systolic),
    bp_diastolic = as.integer(bp_diastolic),
    pulse = as.integer(pulse),
    blood_glucose = as.integer(blood_glucose),
    temperature = as.numeric(temperature),
    vigilance_4ISS = as.factor(vigilance_4ISS),
    gaze_head_deviation_4ISS = as.factor(gaze_head_deviation_4ISS),
    hemiparesis_4ISS = as.factor(hemiparesis_4ISS),
    aphasia_dysarthria_4ISS = as.factor(aphasia_dysarthria_4ISS),
    score_4ISS = as.integer(score_4ISS),
    GCS_eyes = as.integer(GCS_eyes),
    GCS_motor = as.integer(GCS_motor),
    GCS_verbal = as.integer(GCS_verbal),
    score_GCS = as.integer(score_GCS),
    infection = as.factor(infection),
    NIHSS_at_admission = as.integer(NIHSS_at_admission),
    medication_anticoags = as.factor(medication_anticoags),
    medication_antiplatelets = as.factor(medication_antiplatelets),
    TOO = as.numeric(TOO),
    exclusion_TOO = as.factor(exclusion_TOO),
    exclusion_anticoag_status = as.factor(exclusion_anticoag_status),
    exclusion_surgery_status = as.factor(exclusion_surgery_status),
    DTN = as.integer(DTN),
    DTCT = as.integer(DTCT),
    diagnosis_admission = as.character(diagnosis_admission),
    diagnosis_discharge = as.character(diagnosis_discharge),
    IVT = as.factor(IVT),
    EVT = as.factor(EVT),
    mortality = as.factor(mortality),
    pupil_reaction_abnormal = as.factor(pupil_reaction_abnormal),
    consciousness_impaired = as.factor(consciousness_impaired),
    pre_existing_neurological_deficit = as.factor(pre_existing_neurological_deficit),
    ways_of_admission_transport = as.factor(ways_of_admission_transport),
    ways_of_admission_origin = as.factor(ways_of_admission_origin),
    year_of_arrival = as.factor(year_of_arrival),
    diagnosis_discharge = as.factor(diagnosis_discharge)
  )

```

**Glimpse at the data**

```{r}

# Display the first few rows of the dataset
str(matched_data)

```

## Unmatched, Matched and Unprocessed, Train and Testdata

### Get Raw Data

**Extract just the variables available in the Dataset after preprocessing**

```{r}

matched_data <- matched_data %>%
  select(all_of(colnames(matched_data)))

```

### Get Train and Test-Data

Same Train- and Test-Data split as for model training

```{r}

# Same seed as for the Splits for the model training
set.seed(456)

# Split the dataset into training and testing sets
matched_data_split <- initial_split(matched_data,
                                         prop = 0.8,
                                         strata = EVT) # ensures that EVT prevalence is similar in train and testdata


# create seperate dataframes
train_data <- matched_data_split %>% training()
test_data <- matched_data_split %>% testing()

```

### Impute Data

**Select Variables for Imputation**

Select the identical set of variables used in the machine learning scripts. Replicating this selection is essential to show the correct data characteristics across different dataframes in different steps in the ML pipeline.

```{r}

# select imputation variables in the training data
train_data <- train_data %>%
  select(
    gender,
    age, 
    bp_systolic,
    bp_diastolic,
    pulse,
    blood_glucose,
    temperature,
    vigilance_4ISS,
    gaze_head_deviation_4ISS,
    hemiparesis_4ISS, 
    aphasia_dysarthria_4ISS,
    score_4ISS,
    GCS_eyes,
    GCS_motor,
    GCS_verbal,
    score_GCS,
    NIHSS_at_admission,
    pupil_reaction_abnormal, 
    consciousness_impaired,
    pre_existing_neurological_deficit,
    DTN,
    DTCT,
    year_of_arrival,
    ways_of_admission_transport,
    ways_of_admission_origin,
    medication_anticoags,
    medication_antiplatelets,
    TOO,
    exclusion_TOO,
    exclusion_anticoag_status,
    diagnosis_discharge,
    IVT,
    EVT
  )

# select same imputation variables in the test data
test_data <- test_data %>%
  select(
    gender,
    age, 
    bp_systolic,
    bp_diastolic,
    pulse,
    blood_glucose,
    temperature,
    vigilance_4ISS,
    gaze_head_deviation_4ISS,
    hemiparesis_4ISS, 
    aphasia_dysarthria_4ISS,
    score_4ISS,
    GCS_eyes,
    GCS_motor,
    GCS_verbal,
    score_GCS,
    NIHSS_at_admission,
    pupil_reaction_abnormal, 
    consciousness_impaired,
    pre_existing_neurological_deficit,
    DTN,
    DTCT,
    year_of_arrival,
    ways_of_admission_transport,
    ways_of_admission_origin,
    medication_anticoags,
    medication_antiplatelets,
    TOO,
    exclusion_TOO,
    exclusion_anticoag_status,
    diagnosis_discharge,
    IVT,
    EVT
  )


```

**Set Data Types for Imputation**

```{r}

# Ensure correct data types for train_data
train_data <- train_data %>%
  mutate(
    # Ordinal categorical variables
    vigilance_4ISS = factor(vigilance_4ISS,
                            levels = c(0, 1, 2),
                            ordered = TRUE),
    gaze_head_deviation_4ISS = factor(gaze_head_deviation_4ISS,
                                      levels = c(0, 1, 2),
                                      ordered = TRUE),
    hemiparesis_4ISS = factor(hemiparesis_4ISS,
                              levels = c(0, 1, 2),
                              ordered = TRUE),
    
    # Categorical (non-ordinal) variables
    aphasia_dysarthria_4ISS = factor(aphasia_dysarthria_4ISS),
    gender = factor(gender),
    medication_anticoags = factor(medication_anticoags),
    exclusion_TOO = as_factor(exclusion_TOO),
    pupil_reaction_abnormal = factor(pupil_reaction_abnormal),
    exclusion_anticoag_status = factor(exclusion_anticoag_status),
    consciousness_impaired = factor(consciousness_impaired),
    ways_of_admission_transport = factor(ways_of_admission_transport),
    ways_of_admission_origin = factor(ways_of_admission_origin),
    year_of_arrival = factor(year_of_arrival),
    diagnosis_discharge = factor(diagnosis_discharge),

    
    # Outcome variable
    EVT = factor(EVT, levels = c("0", "1")),
    
    # Convert IVT to factor
    IVT = factor(IVT)
  )

# Apply the same transformations to test_data
test_data <- test_data %>%
  mutate(
    # Ordinal categorical variables (0,1,2)
    vigilance_4ISS = factor(vigilance_4ISS,
                            levels = c(0, 1, 2),
                            ordered = TRUE),
    gaze_head_deviation_4ISS = factor(gaze_head_deviation_4ISS,
                                      levels = c(0, 1, 2),
                                      ordered = TRUE),
    hemiparesis_4ISS = factor(hemiparesis_4ISS,
                              levels = c(0, 1, 2),
                              ordered = TRUE),
    aphasia_dysarthria_4ISS = factor(aphasia_dysarthria_4ISS),
    gender = factor(gender),
    medication_anticoags = factor(medication_anticoags),
    exclusion_TOO = factor(exclusion_TOO),
    pupil_reaction_abnormal = factor(pupil_reaction_abnormal),
    exclusion_anticoag_status = factor(exclusion_anticoag_status),
    consciousness_impaired = factor(consciousness_impaired),
    ways_of_admission_transport = factor(ways_of_admission_transport),
    ways_of_admission_origin = factor(ways_of_admission_origin),
    year_of_arrival = factor(year_of_arrival),
    
    EVT = factor(EVT, levels = c("0", "1")),
    
    IVT = factor(IVT)
  )

```

We analyzed the `data_manipulation_5_imputation.qmd` file and found that `step_impute_bag()` provides the most suitable imputation method for our dataset.

Each feature with missing values is treated as the **target (label)**, while all other available features act as **predictors**. Multiple decision trees (like in Random Forest, but each tree uses all predictors) are trained on different random subsets of the data. The missing values are then predicted using these models, and this process is repeated separately for each variable with missing data.

```{r}

# Define the recipe for imputation using bagged trees (if needed)
imputation_recipe <- recipe(EVT ~ ., data = train_data) %>%
  step_impute_bag(all_predictors())

# Train the recipe using only the training data
trained_imputation_recipe <- imputation_recipe %>%
  prep(training = train_data, retain = TRUE)

# Apply the trained recipe separately to training and test data (if imputation is required)
train_data <- bake(trained_imputation_recipe, new_data = train_data)
test_data  <- bake(trained_imputation_recipe, new_data = test_data)

```

### Create Four Dataframes for Comparison

Add a column "Dataset" to each dataframe to distinguish data sources for later grouping

```{r}

unmatched_data <- unmatched_data %>% mutate(Dataset = "Unmatched")

matched_data <- matched_data %>% mutate(Dataset = "Matched")

train_data <-train_data %>% mutate(Dataset = "Training")

test_data <- test_data %>% mutate(Dataset = "Test")

```

### Calculate the Sum of the 4I-SS

This step must be performed after imputation to ensure that, in the rare cases where one of the four items is NA, the sum calculation does not produce an error.

```{r}

# Unmacthed data
unmatched_data <- unmatched_data %>% 
  mutate(
    sum_4ISS = factor(
      as.numeric(as.character(vigilance_4ISS)) +
        as.numeric(as.character(gaze_head_deviation_4ISS)) +
        as.numeric(as.character(hemiparesis_4ISS)) +
        as.numeric(as.character(aphasia_dysarthria_4ISS)),
      ordered = TRUE
    ))

# Matched data
matched_data <- matched_data %>% 
  mutate(
    sum_4ISS = factor(
      as.numeric(as.character(vigilance_4ISS)) +
        as.numeric(as.character(gaze_head_deviation_4ISS)) +
        as.numeric(as.character(hemiparesis_4ISS)) +
        as.numeric(as.character(aphasia_dysarthria_4ISS)),
      ordered = TRUE
    ))

# Training data
train_data <- train_data %>% 
  mutate(
    sum_4ISS = factor(
      as.numeric(as.character(vigilance_4ISS)) +
        as.numeric(as.character(gaze_head_deviation_4ISS)) +
        as.numeric(as.character(hemiparesis_4ISS)) +
        as.numeric(as.character(aphasia_dysarthria_4ISS)),
      ordered = TRUE
    ))

# Test data
test_data <- test_data %>% 
  mutate(
    sum_4ISS = factor(
      as.numeric(as.character(vigilance_4ISS)) +
        as.numeric(as.character(gaze_head_deviation_4ISS)) +
        as.numeric(as.character(hemiparesis_4ISS)) +
        as.numeric(as.character(aphasia_dysarthria_4ISS)),
      ordered = TRUE
    ))

```

### Calculate the NIHSS Groups for all Datasets

```{r}

# Unmatched data
unmatched_data <- unmatched_data %>% 
  mutate(
    NIHSS_group = case_when(
      NIHSS_at_admission <= 4              ~ "0-4",
      NIHSS_at_admission > 4 & NIHSS_at_admission <= 15 ~ "5-15",
      NIHSS_at_admission > 15              ~ ">15",
      TRUE                                 ~ NA_character_
    ),
    NIHSS_group = factor(
      NIHSS_group,
      levels  = c("0-4", "5-15", ">15"),
      ordered = TRUE
    )
  )

# Matched and unprocessed data
matched_data <- matched_data %>% 
  mutate(
    NIHSS_group = case_when(
      NIHSS_at_admission <= 4              ~ "0-4",
      NIHSS_at_admission > 4 & NIHSS_at_admission <= 15 ~ "5-15",
      NIHSS_at_admission > 15              ~ ">15",
      TRUE                                 ~ NA_character_
    ),
    NIHSS_group = factor(
      NIHSS_group,
      levels  = c("0-4", "5-15", ">15"),
      ordered = TRUE
    )
  )

# Training data
train_data <- train_data %>% 
  mutate(
    NIHSS_group = case_when(
      NIHSS_at_admission <= 4              ~ "0-4",
      NIHSS_at_admission > 4 & NIHSS_at_admission <= 15 ~ "5-15",
      NIHSS_at_admission > 15              ~ ">15",
      TRUE                                 ~ NA_character_
    ),
    NIHSS_group = factor(
      NIHSS_group,
      levels  = c("0-4", "5-15", ">15"),
      ordered = TRUE
    )
  )

# Test data
test_data <- test_data %>% 
  mutate(
    NIHSS_group = case_when(
      NIHSS_at_admission <= 4              ~ "0-4",
      NIHSS_at_admission > 4 & NIHSS_at_admission <= 15 ~ "5-15",
      NIHSS_at_admission > 15              ~ ">15",
      TRUE                                 ~ NA_character_
    ),
    NIHSS_group = factor(
      NIHSS_group,
      levels  = c("0-4", "5-15", ">15"),
      ordered = TRUE
    )
  )

```

## Select 14 Features from Feature Selection Process + More Selected Variables

We selected the 14 features that were identified during the feature selection step of the training process. These are the only features that are meaningful for comparison, as they were consistently used to train the models. In addition, we included the 4I-SS sum, NIHSS at admission, year of arrival, the mode of transport (EMS with or without an emergency physician), and the origin of admission (EMS via emergency call, drip-and-ship, general practitioner) to verify whether the datasets used in the machine learning pipeline still represent the original data.

Since the selection is repeated across four different iterations, we define a reusable function named `select_features()` to ensure consistency and reproducibility.

We also add the technical `Dataset` column for using the `by =` function later.

**Create Feature Selection Function**

```{r}

select_features <- function(data) {
  dplyr::select(
    data,
    gender,
    age,
    bp_systolic,
    bp_diastolic,
    pulse,
    blood_glucose,
    temperature,
    vigilance_4ISS,
    gaze_head_deviation_4ISS,
    hemiparesis_4ISS,
    aphasia_dysarthria_4ISS,
    sum_4ISS,
    NIHSS_group,
#    medication_anticoags,
    TOO,
#    exclusion_TOO,
#    pupil_reaction_abnormal,
    IVT,
    EVT,
    ways_of_admission_transport,
    ways_of_admission_origin,
    year_of_arrival,
    diagnosis_discharge,
    Dataset
  )
}

```

**Apply Function**

```{r}

unmatched_data <-         select_features(unmatched_data)

matched_data <-           select_features(matched_data)

train_data <-             select_features(train_data)

test_data <-              select_features(test_data)

```

Also set variable types of the `unmatched_data` and `matched_data` for using `bind_rows`.

```{r}

# Apply transformations to matched_data
unmatched_data <- unmatched_data %>%
  mutate(
    gender = factor(gender),
    vigilance_4ISS = factor(vigilance_4ISS,
                            levels = c(0, 1, 2),
                            ordered = TRUE),
    gaze_head_deviation_4ISS = factor(gaze_head_deviation_4ISS,
                                      levels = c(0, 1, 2),
                                      ordered = TRUE),
    hemiparesis_4ISS = factor(hemiparesis_4ISS,
                              levels = c(0, 1, 2),
                              ordered = TRUE),
    aphasia_dysarthria_4ISS = factor(aphasia_dysarthria_4ISS),
    sum_4ISS = as.factor(sum_4ISS),
    EVT = factor(EVT, levels = c("0", "1")),
  )


# Apply transformations to matched_data
matched_data <- matched_data %>%
  mutate(
    gender = factor(gender),
    vigilance_4ISS = factor(vigilance_4ISS,
                            levels = c(0, 1, 2),
                            ordered = TRUE),
    gaze_head_deviation_4ISS = factor(gaze_head_deviation_4ISS,
                                      levels = c(0, 1, 2),
                                      ordered = TRUE),
    hemiparesis_4ISS = factor(hemiparesis_4ISS,
                              levels = c(0, 1, 2),
                              ordered = TRUE),
    aphasia_dysarthria_4ISS = factor(aphasia_dysarthria_4ISS),
    sum_4ISS = as.factor(sum_4ISS),
    EVT = factor(EVT, levels = c("0", "1")),
  )

```

### Combine All Datasets for Comparison

```{r}

# Combine all datasets into one dataframe
combined_data <- bind_rows(unmatched_data, matched_data, train_data, test_data)

```

```{r}

# Set the Dataset column as a factor with the desired order
combined_data <- combined_data %>%
  mutate(Dataset = factor(Dataset, levels = c("Unmatched",
                                              "Matched",
                                              "Training",
                                              "Test")))

```

## Create Crosstable

**Define Labels**

```{r}

# Define the variable labels in a dataframe
variable_labels <- data.frame(
  name = c("gender",
           "age",
           "bp_systolic",
           "bp_diastolic",
           "pulse",
           "blood_glucose",
           "temperature",
           "vigilance_4ISS",
           "gaze_head_deviation_4ISS",
           "hemiparesis_4ISS", 
           "aphasia_dysarthria_4ISS",
           "sum_4ISS",
           "NIHSS_group",
           "medication_anticoags",
           "medication_antiplatelets", 
           "TOO",
           "exclusion_TOO",
           "exclusion_anticoag_status",
           "exclusion_surgery_status", 
           "DTN",
           "DTCT",
           "diagnosis_discharge",
           "pre_existing_neurological_deficit",
           "IVT", 
           "EVT",
           "pupil_reaction_abnormal",
           "consciousness_impaired",
           "mortality",
           "infection",
           "ways_of_admission_transport",
           "ways_of_admission_origin",
           "year_of_arrival"),
  
  label = c("Gender",
            "Age",
            "Systolic BP",
            "Diastolic BP",
            "Pulse",
            "Blood Glucose", 
            "Temperature",
            "4I-SS Vigilance",
            "4I-SS Gaze Head Deviation",
            "4I-SS Hemiparesis", 
            "4I-SS Aphasia Dysarthria",
            "4I-SS",
            "NIHSS at Admission",
            "Medication Anticoags",
            "Medication Antiplatelets", 
            "Time of Onset (TOO)",
            "Exclusion TOO",
            "Exclusion Anticoag Status", 
            "Exclusion Surgery Status",
            "Door to Needle",
            "Door to CT",
            "Diagnosis at Discharge", 
            "Pre-existing Neurological Deficit",
            "IVT",
            "EVT",
            "Abnormal Pupil Reaction", 
            "Consciousness Impaired",
            "Mortality",
            "Infection",
            "Way of Transport",
            "Origin of Transport",
            "Year of Record")
)

# Import labels to combined_data
combined_data <- combined_data %>% 
  import_labels(variable_labels, name_from = "name", label_from = "label")

```

### Crosstable for All Features

**Create Crosstable as html using `crosstable::crosstable()`**

```{r}

# Create the crosstable with min/max and median/IQR only, without the mean
cross_table_all_features <- crosstable::crosstable(combined_data,
                          by = "Dataset",
                          total = "col",
                          percent_pattern = "{n} ({p_col})",
                          percent_digits = 0) %>%
  as_flextable()

# Display the flextable
cross_table_all_features

```

**Safe as html file in downloads**

```{r}

# # Export in Downloads
#  save_as_html(cross_table_all_features, path = "~/Downloads/cross_table_all_features.html")

```

### Crosstable for Selected Features

**Select Features**

```{r}

combined_data_selected_features <- combined_data %>% 
  select(
    Dataset,
    gender,
    age,
    bp_systolic,
    bp_diastolic,
    NIHSS_group,
    sum_4ISS,
    year_of_arrival,
    ways_of_admission_transport,
    ways_of_admission_origin,
    diagnosis_discharge
  )

```

**Create Crosstable as html using `crosstable::crosstable()`**

```{r}

# Create the crosstable with min/max and median/IQR only, without the mean
cross_table_selected_features <- crosstable::crosstable(combined_data_selected_features,
                          by = "Dataset",
                          total = "col",
                          percent_pattern = "{n} ({p_col})",
                          percent_digits = 0) %>%
  as_flextable()

# Display the flextable
cross_table_selected_features

```

**Safe as html file in downloads**

```{r}

# Export in Downloads
save_as_html(cross_table_selected_features,
             path = "~/Downloads/cross_table_selected_features.html")

```

## Sampling Ratio

Pie charts show the changes in class distribution caused by the sampling process. The charts visualize the ratio of EVT (minority class) and No EVT (majority class) both before and after the sampling methods applied by each model.

```{r}

# Original data
total_instances <- 922
minority_count <- 58
majority_count <- total_instances - minority_count

# Sampling parameters for each algorithm
sampling_parameters <- tibble(
  Algorithm = c("Original Data", "OneR (Down)", "kNN (Over and Down)", "Decision Tree (Down)", "Random Forest (Down)", "Neural Network (Over and Down)"),
  Under_Ratio = c(NA, 1.0, 0.6, 1.0, 1.0, 1.0), # NA for Original Data
  Over_Ratio = c(NA, NA, 0.3, NA, NA, 0.5)      # NA where no oversampling is applied
)

# Function to calculate oversampling and downsampling
calculate_sampling <- function(under_ratio, over_ratio, minority, majority) {
  if (!is.na(over_ratio)) {
    # Apply oversampling: Increase minority class
    minority <- round(majority * over_ratio)
  }
  if (!is.na(under_ratio)) {
    # Apply undersampling: Reduce majority class
    majority <- round(minority * under_ratio)
  }
  return(c(minority, majority))
}

# Apply sampling to each algorithm
sampling_results <- sampling_parameters %>%
  rowwise() %>%
  mutate(
    Counts = list(calculate_sampling(Under_Ratio, Over_Ratio, minority_count, majority_count)),
    Minority = Counts[1],
    Majority = Counts[2]
  ) %>%
  select(-Counts)

# Reshape data for plotting
plot_data <- sampling_results %>%
  pivot_longer(cols = c(Minority, Majority), 
               names_to = "Class", 
               values_to = "Count") %>%
  group_by(Algorithm) %>%
  mutate(Percentage = Count / sum(Count) * 100)

# Colors for the classes
colors <- c("Minority" = "#EA5B57FF", "Majority" = "#1B80ADFF")

# Custom legend labels
legend_labels <- c("Majority" = "No EVT", "Minority" = "EVT")

# Create pie charts for each algorithm
plots <- list()

for (algo in unique(plot_data$Algorithm)) {
  data <- plot_data %>% filter(Algorithm == algo)
  
  p <- ggplot(data, aes(x = "", y = Count, fill = Class)) +
    geom_bar(stat = "identity", width = 1, color = "white") +
    coord_polar(theta = "y") +
    labs(
      title = algo,
      subtitle = paste0("Majority: ", data$Count[data$Class == "Majority"],
                         " | Minority: ", data$Count[data$Class == "Minority"])
    ) +
    scale_fill_manual(
      values = colors,
      labels = legend_labels # Custom legend labels
    ) +
    theme_void() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 12),
      plot.subtitle = element_text(hjust = 0.5, size = 10),
      legend.title = element_blank() # Remove "Class" label
    )
  
  plots[[algo]] <- p
}

# Display all pie charts
do.call(grid.arrange, c(plots, ncol = 2))

```
