---
title: "Evaluation"
editor: visual
---

### Load Packages

```{r}

# for rendering pdf files
library(showtext) 

# Project-relative paths & file management
library(here)

library(tidyverse)
library(tidymodels)

# Evaluation
library(yardstick) 
library(PRROC) # calculate both the ROCAUC and PRAUC
library(pROC)  # plot and compare the (Ensemble) ROC Curves

```

### Load Data

```{r}

stroke_triage_data <- read_csv(here("Code",
                                    "Data Preprocessing",
                                    "data_manipulation_6_SFS",
                                    "data_manipulation_6_SFS.csv"))

```

### Load Workflows

```{r}

# Load each workflow directly with here()

final_knn_workflow  <- readRDS(here("Code",
                                    "Machine Learning",
                                    "models",
                                    "final_knn_workflow.rds"))

final_cart_workflow <- readRDS(here("Code",
                                    "Machine Learning",
                                    "models",
                                    "final_cart_workflow.rds"))

final_xgb_workflow  <- readRDS(here("Code",
                                    "Machine Learning",
                                    "models",
                                    "final_xgb_workflow.rds"))

final_rf_workflow   <- readRDS(here("Code",
                                    "Machine Learning",
                                    "models",
                                    "final_rf_workflow.rds"))

final_nn_workflow   <- readRDS(here("Code",
                                    "Machine Learning",
                                    "models",
                                    "final_nn_workflow.rds"))

```

### Load Models

```{r}

# Load each model directly with here()

final_OneR_model <- readRDS(here("Code",
                                 "Machine Learning",
                                 "models",
                                 "final_OneR_model.rds"))

final_knn_model  <- readRDS(here("Code",
                                 "Machine Learning",
                                 "models",
                                 "final_knn_model.rds"))

final_cart_model <- readRDS(here("Code",
                                 "Machine Learning",
                                 "models",
                                 "final_cart_model.rds"))

final_xgb_model  <- readRDS(here("Code",
                                 "Machine Learning",
                                 "models",
                                 "final_xgb_model.rds"))

final_rf_model   <- readRDS(here("Code",
                                 "Machine Learning",
                                 "models",
                                 "final_rf_model.rds"))

final_nn_model   <- readRDS(here("Code",
                                 "Machine Learning",
                                 "models",
                                 "final_nn_model.rds"))

```

\newpage

# Preprocessing

## Train-Test Data Split

```{r}

# Set seed for reproducible and comparable Train-Test split
# across different algorithms
set.seed(456)

# Split the dataset into training and testing sets
stroke_trage_data_split <- initial_split(stroke_triage_data,
                                         prop = 0.8,
                                         strata = EVT) 

# create seperate dataframes
train_data <- stroke_trage_data_split %>% training()
test_data <- stroke_trage_data_split %>% testing()

stroke_trage_data_split

```

## Imputation

Set Data Types for Imputation

```{r}

# Ensure correct data types for train data
train_data <- train_data %>%
  mutate(
    # Ordinal categorical variables (0,1,2)
    vigilance_4ISS = factor(vigilance_4ISS, levels = c(0, 1, 2), ordered = TRUE),
    gaze_head_deviation_4ISS = factor(gaze_head_deviation_4ISS, levels = c(0, 1, 2), ordered = TRUE),
    hemiparesis_4ISS = factor(hemiparesis_4ISS, levels = c(0, 1, 2), ordered = TRUE),

    # Categorical (non-ordinal) variables
    aphasia_dysarthria_4ISS = factor(aphasia_dysarthria_4ISS),
    gender = factor(gender),
    medication_anticoags = factor(medication_anticoags),
    exclusion_TOO = factor(exclusion_TOO),
    pupil_reaction_abnormal = factor(pupil_reaction_abnormal),

    # Ensure outcome variable is a factor
    EVT = factor(EVT, levels = c("0", "1"))  
  )

# Apply the same transformations to test data
test_data <- test_data %>%
  mutate(
    vigilance_4ISS = factor(vigilance_4ISS, levels = c(0, 1, 2), ordered = TRUE),
    gaze_head_deviation_4ISS = factor(gaze_head_deviation_4ISS, levels = c(0, 1, 2), ordered = TRUE),
    hemiparesis_4ISS = factor(hemiparesis_4ISS, levels = c(0, 1, 2), ordered = TRUE),

    aphasia_dysarthria_4ISS = factor(aphasia_dysarthria_4ISS),
    gender = factor(gender),
    medication_anticoags = factor(medication_anticoags),
    exclusion_TOO = factor(exclusion_TOO),
    pupil_reaction_abnormal = factor(pupil_reaction_abnormal),

    EVT = factor(EVT, levels = c("0", "1"))
  )

```

**Apply Bagged Tree Imputation**

We analyzed the `data_manipulation_5_imputation.qmd` file and found that `step_impute_bag()` provides the most suitable imputation method for our dataset.

Each feature with missing values is treated as the **target (label)**, while all other available features act as **predictors**. Multiple decision trees (like in Random Forest, but each tree uses all predictors) are trained on different random subsets of the data. The missing values are then predicted using these models, and this process is repeated separately for each variable with missing data.

```{r}

# Define the recipe for imputation using bagged trees
imputation_recipe <- recipe(EVT ~ ., data = train_data) %>%
  step_impute_bag(all_predictors())  

# Train the recipe using only the training data
trained_imputation_recipe <- imputation_recipe %>%
  prep(training = train_data, retain = TRUE)  

# Apply the trained recipe separately to training and test data
train_data <- bake(trained_imputation_recipe, new_data = train_data)  
test_data  <- bake(trained_imputation_recipe, new_data = test_data)  

```

Convert all Factors to Numeric

```{r}
#| warning: false

train_data <- train_data %>%
  # Convert factors to numeric
  mutate(across(where(is.factor), ~ as.numeric(as.character(.))),
         # The EVT label as a factor
         EVT = factor(EVT, levels = c("1", "0"))) 

test_data <- test_data %>%
  # Convert factors to numeric
  mutate(across(where(is.factor), ~ as.numeric(as.character(.))), 
         # The EVT label as a factor
         EVT = factor(EVT, levels = c("1", "0")))

```

## Select Features Manually

**Get Current Features**

```{r}

colnames(stroke_triage_data)

```

**Select 14 Features from Feature Selection**

```{r}

# for train data
train_data <- train_data %>% 
  select(gender,
         age,
         bp_systolic,
         pulse,
         blood_glucose,
         temperature,
         vigilance_4ISS,
         gaze_head_deviation_4ISS,
         hemiparesis_4ISS,
         aphasia_dysarthria_4ISS,
         medication_anticoags,
         TOO,
         exclusion_TOO,
         pupil_reaction_abnormal,
         EVT)


# for test data
test_data <- test_data %>% 
  select(gender,
         age,
         bp_systolic,
         pulse,
         blood_glucose,
         temperature,
         vigilance_4ISS,
         gaze_head_deviation_4ISS,
         hemiparesis_4ISS,
         aphasia_dysarthria_4ISS,
         medication_anticoags,
         TOO,
         exclusion_TOO,
         pupil_reaction_abnormal,
         EVT)

```

## Set Data Type for CART and Random Forrest

**CART**

```{r}

# Set column classes in test data
test_data_cart <- test_data %>% 
  mutate(
    vigilance_4ISS           = factor(vigilance_4ISS),
    gaze_head_deviation_4ISS = factor(gaze_head_deviation_4ISS),
    hemiparesis_4ISS         = factor(hemiparesis_4ISS),

    # Categorical (non-ordinal) variables
    aphasia_dysarthria_4ISS = factor(aphasia_dysarthria_4ISS),
    gender = factor(gender),
    medication_anticoags = factor(medication_anticoags),
    exclusion_TOO = factor(exclusion_TOO),
    pupil_reaction_abnormal = factor(pupil_reaction_abnormal),

    # Ensure outcome variable is a factor
    EVT = factor(EVT, levels = c("1", "0"))
  )

```

**Random Forrest**

```{r}

test_data_rf <- test_data %>% 
  mutate(
    vigilance_4ISS           = factor(vigilance_4ISS),
    gaze_head_deviation_4ISS = factor(gaze_head_deviation_4ISS),
    hemiparesis_4ISS         = factor(hemiparesis_4ISS),

    # Categorical (non-ordinal) variables
    aphasia_dysarthria_4ISS = factor(aphasia_dysarthria_4ISS),
    gender = factor(gender),
    medication_anticoags = factor(medication_anticoags),
    exclusion_TOO = factor(exclusion_TOO),
    pupil_reaction_abnormal = factor(pupil_reaction_abnormal),

    # Ensure outcome variable is a factor
    EVT = factor(EVT, levels = c("1", "0"))
  )

```

## Create 4I-SS Reference

Create a variable with the sum 4I-SS scores of the training set

```{r}

# Calculate 4I-SS score
score_4ISS_train <- train_data$hemiparesis_4ISS +
                    train_data$vigilance_4ISS +
                    train_data$gaze_head_deviation_4ISS +
                    train_data$aphasia_dysarthria_4ISS

```

The ML model should outperform the 4-Item Stroke Scale (4I-SS). Therefore, the decisions made by the 4I-SS on the same test set need to be calculated and saved in a separate dataframe for comparison.

```{r}

classification_4ISS <- test_data %>% 
  mutate(score_4ISS = hemiparesis_4ISS +
           vigilance_4ISS +
           gaze_head_deviation_4ISS +
           aphasia_dysarthria_4ISS,
         classification_EVT = as_factor(if_else(score_4ISS > 2, 1, 0))) %>% 
  select(score_4ISS, classification_EVT)

```

## Safe the Labels in a Variable

```{r}

# Training labels
label_train <- train_data$EVT

# Test labels
label_test <- test_data$EVT

```

## Safe the 4I-SS scores in a Variable

```{r}

# 4I-SS scores of train data
score_4ISS_train <- score_4ISS_train

# 4I-SS scores of test data
score_4ISS_test  <- classification_4ISS$score_4ISS

```

## Predictions on Train Data (for Calibration)

To reduce overfitting and avoid information leakage, the calibration model (GLM) should be fitted on predictions from the training data rather than on the test data. Ideally, calibration would be performed on an independent validation set. However, the available data is not enough to split the data into a third fully separate dataset.

**Set Data Types**

CART and Random Forrest need some factoral data types for prediction

```{r}

# CART 
train_data_cart <- train_data %>%
  mutate(
    vigilance_4ISS = factor(vigilance_4ISS),
    gaze_head_deviation_4ISS = factor(gaze_head_deviation_4ISS),
    hemiparesis_4ISS = factor(hemiparesis_4ISS),
    aphasia_dysarthria_4ISS = factor(aphasia_dysarthria_4ISS),
    gender = factor(gender),
    medication_anticoags = factor(medication_anticoags),
    exclusion_TOO = factor(exclusion_TOO),
    pupil_reaction_abnormal = factor(pupil_reaction_abnormal),
    EVT = factor(EVT, levels = c("1", "0"))
  )

# Random Forest 
train_data_rf <- train_data %>%
  mutate(
    vigilance_4ISS = factor(vigilance_4ISS),
    gaze_head_deviation_4ISS = factor(gaze_head_deviation_4ISS),
    hemiparesis_4ISS = factor(hemiparesis_4ISS),
    aphasia_dysarthria_4ISS = factor(aphasia_dysarthria_4ISS),
    gender = factor(gender),
    medication_anticoags = factor(medication_anticoags),
    exclusion_TOO = factor(exclusion_TOO),
    pupil_reaction_abnormal = factor(pupil_reaction_abnormal),
    EVT = factor(EVT, levels = c("1", "0"))
  )

```

**Predictions on Train Data 4I-SS**

Clinical scores are primarily designed for classification (risk stratification) rather than for producing well-calibrated outcome probabilities. Since both the Brier score and decision curve analysis require probabilities, we proceed in two stepsilon: (1) transform the 4I-SS to a 0–1 scale using min–max normalization as an initial probability proxy; (2) calibrate these proxy probabilities to better reflect observed outcome frequencies.

```{r}

# Transform the 4I-SS to a 0–1 scale using min–max normalization
probabilities_score_4ISS_train <- (score_4ISS_train - min(score_4ISS_train)) / 
                                  (max(score_4ISS_train) - min(score_4ISS_train))

```

**Predictions on Train Data ML Models**

```{r}

# KNN
probabilities_knn_train <- predict(final_knn_model, 
                                   new_data = train_data, 
                                   type = "prob")$.pred_1

# CART
probabilities_cart_train <- predict(final_cart_model, 
                                    new_data = train_data_cart, 
                                    type = "prob")$.pred_1

# XGBoost
probabilities_xgb_train <- predict(final_xgb_model, 
                                   new_data = train_data, 
                                   type = "prob")$.pred_1

# Random Forest
probabilities_rf_train <- predict(final_rf_model, 
                                  new_data = train_data_rf, 
                                  type = "prob")$.pred_1

# Neural Network
probabilities_nn_train <- predict(final_nn_model, 
                                  new_data = train_data, 
                                  type = "prob")$.pred_1


```

## Predictions on Test Data (for Evaluation)

**Predicitons on Test Data OneR**

The One rule from OneR needs to be manually applied to the data, as `predict()` does not work with the model.

```{r}

# predict with hemiparesis rule
OneR_predictions_test <- as_factor(ifelse(test_data$hemiparesis_4ISS == 2, 1, 0))

# extract test label
test_label <- as_factor(test_data$EVT)

# convert to numerical data to mimic probablities for ROC
probabilities_OneR_test <- as.numeric(as.character(OneR_predictions_test))

```

**Predictions on Test Data 4I-SS**

Clinical scores are primarily designed for classification (risk stratification) rather than for producing well-calibrated outcome probabilities. Since both the Brier score and decision curve analysis require probabilities, we proceed in two stepsilon: (1) transform the 4I-SS to a 0–1 scale using min–max normalization as an initial probability proxy; (2) calibrate these proxy probabilities to better reflect observed outcome frequencies.

```{r}

# Transform the 4I-SS to a 0–1 scale using min–max normalization
probabilities_score_4ISS_test <- (score_4ISS_test - min(score_4ISS_test)) /                                         (max(score_4ISS_test) - min(score_4ISS_test))

```

**Predictions on Test Data Other Models**

```{r}
#| warning: false

# KNN
probabilities_knn_test <- predict(final_knn_model,
                             new_data = test_data,
                             type = "prob")$.pred_1

# CART
probabilities_cart_test <- predict(final_cart_model,
                              new_data = test_data_cart,
                              type = "prob")$.pred_1

# XGBoost
probabilities_xgb_test <- predict(final_xgb_model,
                             new_data = test_data,
                             type = "prob")$.pred_1

# Random Forest
probabilities_rf_test <- predict(final_rf_model,
                            new_data = test_data_rf,
                            type = "prob")$.pred_1

# Neural Network
probabilities_nn_test <- predict(final_nn_model,
                            new_data = test_data,
                            type = "prob")$.pred_1

```

## Store the Predicted Probabilities

We need to use these probabilities later for ROC, PR, Calibration and DCA.

**Store Train Probabilities**

```{r}
  
# Training probabilities for fitting calibration models
calibration_data_train <- tibble(
    obs         = as.numeric(as.character(label_train)),
    KNN         = probabilities_knn_train,
    CART        = probabilities_cart_train,
    XGB         = probabilities_xgb_train,
    RF          = probabilities_rf_train,
    NN          = probabilities_nn_train,
    score_4ISS  = probabilities_score_4ISS_train
  )

```

**Store Test Probabilities**

**For Calibration (without OneR)**

```{r}
  
# Test probabilities for calibartion
calibration_data_test <- tibble(
    obs         = as.numeric(as.character(label_test)),
    KNN         = probabilities_knn_test,
    CART        = probabilities_cart_test,
    XGB         = probabilities_xgb_test,
    RF          = probabilities_rf_test,
    NN          = probabilities_nn_test,
    score_4ISS  = probabilities_score_4ISS_test
  )

```

**For ROC and PR (without OneR)**

```{r}

# Map each model's scores to its name in `model_test_scores` for easy iteration.

model_test_scores <- list(
  `OneR`  = probabilities_OneR_test,
  `KNN`   = probabilities_knn_test,
  `CART`  = probabilities_cart_test,
  `XGB`   = probabilities_xgb_test,
  `RF`    = probabilities_rf_test,
  `NN`    = probabilities_nn_test,
  `4I-SS` = probabilities_score_4ISS_test
)

labels <- label_test

```

# ROC and PROC

**Used colors for both plots**

```{r}

# Use the colors of the NEJM color palette
# We used alpha 0.9 as a filter, that the colors are not as bright

colors_nejm_alpha90 <- c(
  "#BC3C29E6",
  "#0072B5E6",
  "#E18727E6",
  "#20854EE6",
  "#7876B1E6",
  "#6F99ADE6",
  "#FFDC91E6",
  "#EE4C97E6"
)

```

### ROC Curve

```{r}

# Start PDF device for vector output
pdf(file = "~/Downloads/roc_curve.pdf", width = 6, height = 6)

# Configure plotting layout and text scaling
par(mfrow = c(1, 1),  # single panel
    pty = "s",        # square plotting region
    cex = 1.2)        # global text size scaling

# Initialize empty plot area for ROC curves
plot(NULL, xlim = c(0, 100), ylim = c(0, 100), 
     xlab = "1 - Specificity (%)", 
     ylab = "Sensitivity (%)", 
     main = "")

# Loop through model scores and plot ROC curves
for (i in seq_along(model_test_scores)) {
  roc_curve <- roc.curve(
    scores.class0 = model_test_scores[[i]][labels == 1], 
    scores.class1 = model_test_scores[[i]][labels == 0], 
    curve = TRUE
  )
  
  lines(roc_curve$curve[, 1] * 100,  # X: 1 - specificity (%)
        roc_curve$curve[, 2] * 100,  # Y: sensitivity (%)
        col = colors_nejm_alpha90[i],
        lwd = 3)
}

# diagonal line
abline(0, 1, lty = 2, col = "black", lwd = 2)

# Add legend to distinguish models
legend("bottomright",
       legend = names(model_test_scores),
       col = colors_nejm_alpha90,
       lwd = 3,
       cex = 0.8,
       bty = "n")


# Close PDF device to save the file
dev.off()

```

### Precision-Recall Curve

```{r}

# Start PDF graphics device
pdf(file = "~/Downloads/pr_curve.pdf", width = 6, height = 6)

# Configure plot layout and text scaling
par(mfrow = c(1, 1),  # single plot panel
    pty = "s",        # square plotting region
    cex = 1.0)        # base text size

# Initialize empty plot area for PR curves (axes only)
plot(NULL, xlim = c(0, 100), ylim = c(0, 100), 
     xlab = "Recall (%)", 
     ylab = "Precision (%)", 
     main = "")

# Loop through model scores and draw each PR curve
for (i in seq_along(model_test_scores)) {
  pr_curve <- pr.curve(
    scores.class0 = model_test_scores[[i]][labels == 1], 
    scores.class1 = model_test_scores[[i]][labels == 0], 
    curve = TRUE
  )
  
  # Plot precision-recall curve, scaled to percentages
  lines(pr_curve$curve[, 1] * 100,  # recall (%)
        pr_curve$curve[, 2] * 100,  # precision (%)
        col = colors_nejm_alpha90[i],
        lwd = 3)
}

# prevalence line
prevalence <- mean(labels == 1) * 100
abline(h = prevalence, lty = 2, col = "black", lwd = 2)

# Add legend identifying each model curve
legend("topright",
       legend = names(model_test_scores),
       col = colors_nejm_alpha90,
       lwd = 3,
       cex = 0.8,
       bty = "n")


# Close PDF device to finalize file
dev.off()

```

### Combined Presentation

**Create Combined PDF**

```{r}

# Start PDF device for vector output
pdf(file = "~/Downloads/roc_pr_curve.pdf",
    width = 12,
    height = 6)

# Plot configurations for 1x2 layout
par(mfrow = c(1, 2),
    pty = "s",
    cex = 1.2)

# ROC curves
plot(NULL, xlim = c(0, 100), ylim = c(0, 100), 
     xlab = "1 - Specificity (%)", 
     ylab = "Sensitivity (%)")

for (i in seq_along(model_test_scores)) {
  roc_curve <- roc.curve(
    scores.class0 = model_test_scores[[i]][labels == 1], 
    scores.class1 = model_test_scores[[i]][labels == 0], 
    curve = TRUE
  )
  lines(roc_curve$curve[, 1] * 100, 
        roc_curve$curve[, 2] * 100, 
        col = colors_nejm_alpha90[i], lwd = 3)
}

# Add diagonal random line (black, dashed)
abline(0, 1, lty = 2, col = "black", lwd = 2)

legend("bottomright",
       legend = names(model_test_scores),
       col = colors_nejm_alpha90,
       lwd = 4,
       seg.len = 0.6,
       x.intersp = 0.2,
       cex = 1,
       bty = "n",
       inset = c(0.03, 0))


# PR curves
plot(NULL, xlim = c(0, 100), ylim = c(0, 100), 
     xlab = "Recall (%)", 
     ylab = "Precision (%)")

for (i in seq_along(model_test_scores)) {
  pr_curve <- pr.curve(
    scores.class0 = model_test_scores[[i]][labels == 1], 
    scores.class1 = model_test_scores[[i]][labels == 0], 
    curve = TRUE
  )
  lines(pr_curve$curve[, 1] * 100, 
        pr_curve$curve[, 2] * 100, 
        col = colors_nejm_alpha90[i], lwd = 3)
}

# Add prevalence baseline (black, dashed)
prevalence <- mean(labels == 1) * 100
abline(h = prevalence, lty = 2, col = "black", lwd = 2)

legend("topright",
       legend = names(model_test_scores),
       col = colors_nejm_alpha90,
       lwd = 4,
       seg.len = 0.6,
       x.intersp = 0.2,
       cex = 1,
       bty = "n",
       inset = c(0.03, 0))


# Close PDF device
dev.off()

```

**Create Combined PNG**

```{r}

# Start PNG device for raster output
png(file = "~/Downloads/roc_pr_curve.png",
    width = 12,
    height = 6,
    units = "in",
    res = 300)

# Plot configurations for 1x2 layout
par(mfrow = c(1, 2),
    pty = "s",
    cex = 1.2)

# ROC curves
plot(NULL, xlim = c(0, 100), ylim = c(0, 100), 
     xlab = "1 - Specificity (%)", 
     ylab = "Sensitivity (%)")

for (i in seq_along(model_test_scores)) {
  roc_curve <- roc.curve(
    scores.class0 = model_test_scores[[i]][labels == 1], 
    scores.class1 = model_test_scores[[i]][labels == 0], 
    curve = TRUE
  )
  lines(roc_curve$curve[, 1] * 100, 
        roc_curve$curve[, 2] * 100, 
        col = colors_nejm_alpha90[i], lwd = 3)
}

# Add diagonal random line
abline(0, 1, lty = 2, col = "black", lwd = 2)

legend("bottomright",
       legend = names(model_test_scores),
       col = colors_nejm_alpha90,
       lwd = 4,
       seg.len = 0.6,
       x.intersp = 0.2,
       cex = 1,
       bty = "n",
       inset = c(0.03, 0))

# PR curves
plot(NULL, xlim = c(0, 100), ylim = c(0, 100), 
     xlab = "Recall (%)", 
     ylab = "Precision (%)")

for (i in seq_along(model_test_scores)) {
  pr_curve <- pr.curve(
    scores.class0 = model_test_scores[[i]][labels == 1], 
    scores.class1 = model_test_scores[[i]][labels == 0], 
    curve = TRUE
  )
  lines(pr_curve$curve[, 1] * 100, 
        pr_curve$curve[, 2] * 100, 
        col = colors_nejm_alpha90[i], lwd = 3)
}

# Add prevalence baseline
prevalence <- mean(labels == 1) * 100
abline(h = prevalence, lty = 2, col = "black", lwd = 2)

legend("topright",
       legend = names(model_test_scores),
       col = colors_nejm_alpha90,
       lwd = 4,
       seg.len = 0.6,
       x.intersp = 0.2,
       cex = 1,
       bty = "n",
       inset = c(0.03, 0))

# Close PNG device
dev.off()

```

## Area Under the Curves

```{r}


# Initialize a data frame to store AUC values
auc_table <- data.frame(Model = character(), ROC_AUC = numeric(), PR_AUC = numeric(), stringsAsFactors = FALSE)

# Calculate ROC AUC and PR AUC values for each model
for (model_name in names(model_test_scores)) {
  # ROC AUC
  roc_result <- roc(labels, model_test_scores[[model_name]], levels = c(0, 1), direction = "<")
  roc_auc_value <- as.numeric(auc(roc_result))  # Convert ROC AUC to numeric
  
  # PR AUC
  pr_result <- pr.curve(scores.class0 = model_test_scores[[model_name]][labels == 1], 
                        scores.class1 = model_test_scores[[model_name]][labels == 0], 
                        curve = FALSE)
  pr_auc_value <- pr_result$auc.integral  # Extract PR AUC
  
  # Store the AUC values in the table
  auc_table <- auc_table %>%
    add_row(Model = model_name, ROC_AUC = roc_auc_value, PR_AUC = pr_auc_value)
}

# Display the AUC table
print(auc_table)

```

## Sociodemographic Subgroup Risk Analysis

**Create Datasets**

For OneR, kNN, ANN

```{r}

# Create a copy of train_data for risk analysis and age groups
train_data_risk <- train_data %>%
  mutate(
    age_group = case_when(
      age < 60 ~ "<60",
      age >= 60 & age <= 80 ~ "60-80",
      age > 80 ~ ">80"
    )
  )

```

**Adjust Datatypes**

Random Forrest

```{r}

train_data_risk_rf <- train_data_risk %>% mutate(
    gender = as.factor(gender),
    medication_anticoags = as.factor(medication_anticoags),
    exclusion_TOO = as.factor(exclusion_TOO),
    pupil_reaction_abnormal = as.factor(pupil_reaction_abnormal),
    aphasia_dysarthria_4ISS = as.factor(aphasia_dysarthria_4ISS),
    
    # Ordinal variables as ordered factors
    vigilance_4ISS = factor(vigilance_4ISS),
    gaze_head_deviation_4ISS = factor(gaze_head_deviation_4ISS),
    hemiparesis_4ISS = factor(hemiparesis_4ISS),
    EVT = factor(EVT, levels = c("1", "0"))  # Ensure correct factor levels
)

```

### Generate Model Predictions

Predictions are generated for each model on the test data.

The age column is removed to prevent any influence on classification, as it was not included in the training data.

```{r}
#| message: false

# Generate predictions for each model on the training data with risk groups

# OneR model predictions using a rule on the hemiparesis feature
# Here, a hemiparesis score of 2 is assigned a prediction of 1 (positive), otherwise 0 (negative)
OneR_predictions_train_risk <- factor(ifelse(train_data_risk$hemiparesis_4ISS == 2, 1, 0), levels = c(0, 1))
OneR_probabilities_risk <- as.numeric(as.character(OneR_predictions_train_risk))  # Convert factor predictions to numeric probabilities

# Generate predictions for the KNN model on train_data_risk
# Exclude the age_group variable to ensure it doesn’t influence predictions
KNN_probabilities_risk <- predict(final_knn_model, 
                                  new_data = train_data_risk %>% select(-age_group), 
                                  type = "prob")$.pred_1  # Extract the predicted probability for class "1"


# Generate predictions for the ANN model on train_data_risk
XGB_probabilities_risk <- predict(final_xgb_model, 
                                  new_data = train_data_risk %>% select(-age_group), 
                                  type = "prob")$.pred_1  # Extract probability for class "1"

# Generate predictions for the Random Forest model on train_data_risk_trees
RF_probabilities_risk <- predict(final_rf_model, 
                                 new_data = train_data_risk_rf %>% select(-age_group), 
                                 type = "prob")$.pred_1  # Extract probability for class "1"

# Generate predictions for the ANN model on train_data_risk
ANN_probabilities_risk <- predict(final_nn_model, 
                                  new_data = train_data_risk %>% select(-age_group), 
                                  type = "prob")$.pred_1  # Extract probability for class "1"


# Assign row names to train_data_risk based on a unique identifier or row numbers
# This ensures consistency in indexing when assigning names to each probability prediction
rownames(train_data_risk) <- as.character(1:nrow(train_data_risk))

# Assign names to each model’s probability output using the row names in train_data_risk
# This maintains alignment between the data and model predictions
names(OneR_probabilities_risk) <- rownames(train_data_risk)
names(KNN_probabilities_risk) <- rownames(train_data_risk)
names(XGB_probabilities_risk) <- rownames(train_data_risk)
names(RF_probabilities_risk) <- rownames(train_data_risk)
names(ANN_probabilities_risk) <- rownames(train_data_risk)

# Combine all model probability predictions into a single list
# This structure allows for easy access to each model's predictions when calculating ensemble scores or ROC curves
model_test_scores_risk <- list(
  OneR = OneR_probabilities_risk,
  KNN = KNN_probabilities_risk,
  XGB = XGB_probabilities_risk,
  RF = RF_probabilities_risk,
  ANN = ANN_probabilities_risk
)

```

**Define Group Lists**

Groups for Gender and Age are defined, and colors are assigned to each for plotting.

```{r}

# Split the training data into risk groups by Gender and Age, preserving row names
# This allows us to easily subset the data for each group and keep track of individual rows

risk_groups <- list(
  Gender = split(train_data_risk, train_data_risk$gender),  # Split data into "0" and "1" 
  Age = split(train_data_risk, train_data_risk$age_group)   # Split data into age groups
)

# Set labels

age_group_labels <- c("< 60", "60 - 80", "> 80")  # Age group labels

gender_labels <- c("Female", "Male")  # Gender labels for plotting and interpretation

```

**Verify group sizes**

If groups are to small ROC Curves can be missleading

```{r}

# Check subgroup sizes
for (i in 1:length(risk_groups$Age)) {
  subgroup <- risk_groups$Age[[i]]
  group_size <- nrow(subgroup)
  print(paste("Age Group", age_group_labels[i], "Size:", group_size))
  
  # Check class distribution
  EVT_labels_age <- factor(subgroup$EVT, levels = c(0, 1))
  class_counts <- table(EVT_labels_age)
  print(paste("Class distribution for Age Group", age_group_labels[i], ":"))
  print(class_counts)
}

```

**Aggregate the Predictions of all Models**

```{r}

# Convert the list of model predictions into a data frame
model_test_scores_df <- as.data.frame(model_test_scores_risk)

# Compute the ensemble prediction by averaging the predictions across models
ensemble_predictions <- rowMeans(model_test_scores_df, na.rm = TRUE)

```

**Further Preprocessing**

```{r}

# Extract actual labels from the dataset, ensuring they are in the correct factor format
EVT_labels <- factor(train_data_risk$EVT, levels = c(0, 1))
names(EVT_labels) <- rownames(train_data_risk)  # Assign names to ensure alignment by row

# Verify that the names in EVT_labels match the names in ensemble_predictions
all_names_match <- all(names(EVT_labels) == names(ensemble_predictions))

# Align EVT_labels and ensemble_predictions by matching names if they do not match
if (!all_names_match) {
  common_indices <- intersect(names(EVT_labels), names(ensemble_predictions))  # Find common indices
  EVT_labels <- EVT_labels[common_indices]  # Subset labels to common indices
  ensemble_predictions <- ensemble_predictions[common_indices]  # Subset predictions to common indices
}

# Remove any entries with NA values to ensure clean data for ROC analysis
na_indices <- which(is.na(EVT_labels) | is.na(ensemble_predictions))
if (length(na_indices) > 0) {
  EVT_labels <- EVT_labels[-na_indices]  # Remove NA values from EVT_labels
  ensemble_predictions <- ensemble_predictions[-na_indices]  # Remove NA values from ensemble_predictions
}

```

### Ensemble Stratified ROC Curves and AUC Values

**Used colors for both plots**

```{r}

gender_colors_risk <- c("#FFB900FF", "#5773CCFF")
age_colors_risk <- c("#A1CAF6FF", "#4C6FA1FF", "#1E2F46FF")

```

### ROC Gender

```{r}

# Start PDF device for vector output
pdf(file = "~/Downloads/roc_risk_gender.pdf", width = 6, height = 6)

# Plot settings
par(mfrow = c(1, 1),    # single plot
    pty = "s",          # square plotting region
    cex = 1.0)          # base font scaling

# Gender groups and AUC storage
gender_groups <- c("Female ", "Male     ")  
gender_aucs <- c()  # Empty vector to store AUC values for gender

plot_initialized <- FALSE

# Loop over each gender group, plot ROC, calculate AUC
for (i in 1:length(gender_groups)) {
  gender_code <- as.character(i - 1)  
  indices <- which(train_data_risk$gender == gender_code)
  EVT_labels_subgroup <- EVT_labels[indices]               
  ensemble_predictions_subgroup <- ensemble_predictions[indices]

  if (length(unique(EVT_labels_subgroup)) < 2) next

  ensemble_roc_subgroup <- roc(
    EVT_labels_subgroup,
    ensemble_predictions_subgroup,
    levels = c(0, 1),
    direction = "<",
    percent = TRUE
  )

  gender_aucs[i] <- round(ensemble_roc_subgroup$auc, 2)

  if (!plot_initialized) {
    plot(ensemble_roc_subgroup,
         col = gender_colors_risk[i], lwd = 4,
         legacy.axes = TRUE,
         xlab = "1 - Specificity in %",
         ylab = "Sensitivity in %")
    plot_initialized <- TRUE
  } else {
    plot(ensemble_roc_subgroup,
         col = gender_colors_risk[i], lwd = 4,
         legacy.axes = TRUE, add = TRUE)
  }
}

# Create legend text with AUC values
gender_labels_with_auc <- paste(gender_groups, "(AUC:", gender_aucs, ")")

# Add legend
legend("bottomright",
       legend = gender_labels_with_auc,
       col = gender_colors_risk,
       lwd = 4,
       cex = 0.8,
       bty = "n",
       inset = c(0, 0))

# Close PDF device
dev.off()

```

### ROC Age Groups

```{r}

# Open PDF device for vector output
pdf(file = "~/Downloads/roc_risk_age.pdf", width = 6, height = 6)

# Plot configuration
par(mfrow = c(1, 1),  # single plot panel
    pty = "s",        # square plot region
    cex = 1.0)        # global text scaling

# Age Group ROC Curves 
age_groups <- unique(train_data_risk$age_group)
age_group_labels <- c("     < 60", " 60 - 80", "     > 80")
age_aucs <- c()  # Empty vector to store AUC values

plot_initialized <- FALSE

# Loop through age groups, plot ROC, calculate AUC
for (i in seq_along(age_groups)) {
  age_group <- age_groups[i]
  indices <- which(train_data_risk$age_group == age_group)
  EVT_labels_subgroup <- EVT_labels[indices]                
  ensemble_predictions_subgroup <- ensemble_predictions[indices]

  if (length(unique(EVT_labels_subgroup)) < 2) next

  ensemble_roc_subgroup <- roc(
    EVT_labels_subgroup,
    ensemble_predictions_subgroup,
    levels = c(0, 1),
    direction = "<",
    percent = TRUE
  )

  age_aucs[i] <- round(ensemble_roc_subgroup$auc, 2)

  if (!plot_initialized) {
    plot(ensemble_roc_subgroup,
         col = age_colors_risk[i], lwd = 4,
         legacy.axes = TRUE,
         xlab = "1 - Specificity in %",
         ylab = "Sensitivity in %")
    plot_initialized <- TRUE
  } else {
    plot(ensemble_roc_subgroup,
         col = age_colors_risk[i], lwd = 4,
         legacy.axes = TRUE,
         add = TRUE)
  }
}

# Create legend labels with AUCs
age_labels_with_auc <- paste(age_group_labels, "(AUC:", age_aucs, ")")

# Add legend
legend("bottomright",
       legend = age_labels_with_auc,
       col = age_colors_risk,
       lwd = 4,
       cex = 0.8,
       bty = "n",
       inset = c(0, 0))

# Close PDF device to save
dev.off()

```

### Combined Presentation

**Create Combine Risk PDF**

```{r}

# Open PDF device (safer and fully reproducible)
pdf(file = "~/Downloads/roc_risk.pdf", width = 12, height = 6)

# Configure a 1x2 plotting layout and increase font size globally
par(mfrow = c(1, 2),
    pty = "s",
    cex = 1.2)

# Gender Group ROC Curves
gender_groups <- c("Female ", "Male     ")  
gender_aucs <- c()
plot_initialized <- FALSE

for (i in 1:length(gender_groups)) {
  gender_code <- as.character(i - 1)  
  indices <- which(train_data_risk$gender == gender_code)
  EVT_labels_subgroup <- EVT_labels[indices]               
  ensemble_predictions_subgroup <- ensemble_predictions[indices]
  if (length(unique(EVT_labels_subgroup)) < 2) next

  ensemble_roc_subgroup <- roc(EVT_labels_subgroup, ensemble_predictions_subgroup,
                               levels = c(0, 1), direction = "<", percent = TRUE)
  gender_aucs[i] <- round(ensemble_roc_subgroup$auc, 0)

  if (!plot_initialized) {
    plot(ensemble_roc_subgroup,
         col = gender_colors_risk[i], lwd = 4,
         legacy.axes = TRUE,
         xlab = "1 - Specificity in %",
         ylab = "Sensitivity in %")
    plot_initialized <- TRUE
  } else {
    plot(ensemble_roc_subgroup,
         col = gender_colors_risk[i], lwd = 4,
         legacy.axes = TRUE, add = TRUE)
  }
}

gender_labels_with_auc <- paste(gender_groups, "( AUC:", gender_aucs, ")")
legend("bottomright", legend = gender_labels_with_auc,
       col = gender_colors_risk,
       lwd = 4,
       seg.len = 0.6,
       x.intersp = 0.2,
       bty = "n",
       inset = c(0, 0))


# --- Age Group ROC Curves ---
age_groups <- unique(train_data_risk$age_group)
age_group_labels <- c("    < 60", "60 - 80", "    > 80")
age_aucs <- c()
plot_initialized <- FALSE

for (i in 1:length(age_groups)) {
  age_group <- age_groups[i]
  indices <- which(train_data_risk$age_group == age_group)
  EVT_labels_subgroup <- EVT_labels[indices]                
  ensemble_predictions_subgroup <- ensemble_predictions[indices]
  if (length(unique(EVT_labels_subgroup)) < 2) next

  ensemble_roc_subgroup <- roc(EVT_labels_subgroup, ensemble_predictions_subgroup,
                               levels = c(0, 1), direction = "<", percent = TRUE)
  age_aucs[i] <- round(ensemble_roc_subgroup$auc, 0)

  if (!plot_initialized) {
    plot(ensemble_roc_subgroup,
         col = age_colors_risk[i], lwd = 4,
         legacy.axes = TRUE,
         xlab = "1 - Specificity in %",
         ylab = "Sensitivity in %")
    plot_initialized <- TRUE
  } else {
    plot(ensemble_roc_subgroup,
         col = age_colors_risk[i], lwd = 4,
         legacy.axes = TRUE, add = TRUE)
  }
}

age_labels_with_auc <- paste(age_group_labels, "( AUC:", age_aucs, ")")
legend("bottomright", legend = age_labels_with_auc,
       col = age_colors_risk,
       lwd = 4,
       seg.len = 0.6,
       x.intersp = 0.2,
       bty = "n",
       inset = c(0, 0))

# Reset layout to default
par(mfrow = c(1, 1))

# Close PDF device
dev.off()

```

**Create Combine Risk PNG**

```{r}

png(file = "~/Downloads/roc_risk.png",
    width = 12,
    height = 6,
    units = "in",
    res = 300)

par(mfrow = c(1, 2),
    pty = "s",
    cex = 1.2)

# Gender Group ROC Curves
gender_groups <- c("Female ", "Male     ")
gender_aucs <- c()

# Base R plot frame (same as ROC/PR)
plot(NULL, xlim = c(0, 100), ylim = c(0, 100),
     xlab = "1 - Specificity (%)",
     ylab = "Sensitivity (%)")

for (i in 1:length(gender_groups)) {
  gender_code <- as.character(i - 1)
  indices <- which(train_data_risk$gender == gender_code)
  EVT_labels_subgroup <- EVT_labels[indices]
  ensemble_predictions_subgroup <- ensemble_predictions[indices]
  if (length(unique(EVT_labels_subgroup)) < 2) next

  ensemble_roc_subgroup <- roc(EVT_labels_subgroup, ensemble_predictions_subgroup,
                               levels = c(0, 1), direction = "<", percent = TRUE)
  gender_aucs[i] <- round(ensemble_roc_subgroup$auc, 0)

  # Draw as lines on base plot (like ROC/PR)
  lines(100 - ensemble_roc_subgroup$specificities,
        ensemble_roc_subgroup$sensitivities,
        col = gender_colors_risk[i], lwd = 4)
}

abline(0, 1, lty = 2, col = "black", lwd = 2)

gender_labels_with_auc <- paste(gender_groups, "( AUC:", gender_aucs, ")")
legend("bottomright", legend = gender_labels_with_auc,
       col = gender_colors_risk, lwd = 4,
       seg.len = 0.6, x.intersp = 0.2, bty = "n", inset = c(0, 0))


# --- Age Group ROC Curves ---
age_groups <- unique(train_data_risk$age_group)
age_group_labels <- c("    < 60", "60 - 80", "    > 80")
age_aucs <- c()

plot(NULL, xlim = c(0, 100), ylim = c(0, 100),
     xlab = "1 - Specificity (%)",
     ylab = "Sensitivity (%)")

for (i in 1:length(age_groups)) {
  age_group <- age_groups[i]
  indices <- which(train_data_risk$age_group == age_group)
  EVT_labels_subgroup <- EVT_labels[indices]
  ensemble_predictions_subgroup <- ensemble_predictions[indices]
  if (length(unique(EVT_labels_subgroup)) < 2) next

  ensemble_roc_subgroup <- roc(EVT_labels_subgroup, ensemble_predictions_subgroup,
                               levels = c(0, 1), direction = "<", percent = TRUE)
  age_aucs[i] <- round(ensemble_roc_subgroup$auc, 0)

  lines(100 - ensemble_roc_subgroup$specificities,
        ensemble_roc_subgroup$sensitivities,
        col = age_colors_risk[i], lwd = 4)
}

abline(0, 1, lty = 2, col = "black", lwd = 2)

age_labels_with_auc <- paste(age_group_labels, "( AUC:", age_aucs, ")")
legend("bottomright", legend = age_labels_with_auc,
       col = age_colors_risk, lwd = 4,
       seg.len = 0.6, x.intersp = 0.2, bty = "n", inset = c(0, 0))

par(mfrow = c(1, 1))
dev.off()

```

# Calibration Curves

**Introduction in Scikit-Learn**

> When performing classification one often wants to predict not only the class label, but also the associated probability. This probability gives some kind of confidence on the prediction. This example demonstrates how to visualize how well calibrated the predicted probabilities are using calibration curves, also known as reliability diagrams.

<https://scikit-learn.org/stable/auto_examples/calibration/plot_calibration_curve.html>

## Introduction

**1. Discrimination vs Calibration**

Standard Machine Learning (ML) classification models are primarily designed to **rank** or **separate** classes, not to predict exact probabilities. During model training, hyperparameters are typically optimized for metrics like **ROCAUC**, which measures **discrimination** (how well the model distinguishes between sick and healthy) rather than how "true" the predicted percentages are.

**Example:** If a model predicts a probability of 0.11 for all sick patients and 0.10 for all healthy ones, the **ROCAUC is a perfect 1.0**. However, the probabilities are clinically useless because they don't reflect the actual risk.

**2. Systematic Biases in ML Algorithms**

Different algorithms have known tendencies to either overestimate or underestimate risk. This is often an inherent part of their mathematical structure:

-   **Boosting (e.g., XGBoost):** These models aggressively minimize loss, which often pushes probabilities toward the extremes (0 or 1). This results in **Overconfidence**.

-   **Random Forest:** Because it averages the results of many individual trees, it rarely predicts exactly 0 or 1. This tends to pull predictions toward the center, resulting in **Underconfidence**.

-   **Imbalanced Data:** In datasets with low prevalence (e.g., only 8% LVO cases), models are naturally biased toward low probabilities, causing a **systematic shift** in the output scale.

**3. Why Calibration Matters**

If your goal is only simple classification (e.g., "Is it A or B?"), these distorted probabilities might not matter. However, when working with **probability-based metrics**, calibration becomes mandatory:

-   **Brier Score:** Measures the squared distance between the predicted probability and the actual label (0 or 1).

-   **Decision Curve Analysis (DCA):** The x-axis represents clinical **threshold probabilities**. If the model's probabilities are not "real," the Net Benefit calculation fails because you are comparing clinical preferences to distorted algorithm scores.

**4. How Calibration Works (Platt Scaling)**

To fix this, we use a technique called **Calibration** (often via **Platt Scaling**).

We train a second, simpler model—usually a **Logistic Regression (GLM)**—with a specific setup:

-   **Input (Feature):** Just one Feature, the "raw" predicted probabilities from the ML model.

-   **Target (Label):** The actual outcomes (0 or 1).

This transforms the biased ML output into a **calibrated probability** that can be safely used for clinical decision-making.

**5. Impact of Undersampling and Oversampling**

TRIPOD-AI item 13 says:

> "If class imbalance methods were used, state why and how this was done, and any subsequent methods to recalibrate the model or the model predictions."

and the explanation:

> "Imbalance corrections have an impact on model calibration (van den Goorbergh et al, J Am Med Inform Assoc 2022), yielding probability estimates that are too high \[...\] describe the methods used to recalibrate"

The use of undersampling, oversampling and SMOTE results in poorly calibrated and , regarding the probabities, strongly overestimated models (van den Goorbergh et al, J Am Med Inform Assoc 2022).

<https://doi.org/10.1093/jamia/ocac093>

**We exclude the OneR Model from calibration, because the predicted probabilities are not defined.**

## Fit on Out-of-Fold Predictions

Calibration models tend to overfit if they are trained (i.e., finding the sigmoid fits) on the same dataset on which they should also predict the real probabilities. To prevent the model from overfitting, the calibration models are fitted using the probabilities from the cross-validation folds, identical to the model scripts, using identical models and ML workflows (i.e., pipelines).

Here, we find the parameters of the logistic regression model using the training set with 5-fold cross-validation: the model is fit on the training data and applied to the test data. This 5-fold CV approach gives us a better estimate of whether the calibration generalises well to unseen data (here: predictions), similar to standard ML training practices.

This method is similar to the scikit-learn function `CalibratedClassifierCV` and is also proposed in Niculescu-Mizil and Caruana (2005).\
\
Niculescu-Mizil, Alexandru, and Rich Caruana. "Obtaining Calibrated Probabilities from Boosting." UAI. Vol. 5. 2005.

**Create CV-Folds**

```{r}

set.seed(456)
cv_folds <- vfold_cv(train_data,
                     v = 5,
                     strata = EVT)

```

**Function for Out-of-Fold (OOF) Predictions**

```{r}

get_oof_predictions <- function(workflow, folds, train_data_full, 
                                 prep_function = NULL) {
  
 map_dfr(1:nrow(folds), function(i) {
    
    split <- folds$splits[[i]]
    train_fold <- training(split)
    validation_fold <- testing(split)
    
    # Convert to factors for CART/RF
    if (!is.null(prep_function)) {
      train_fold <- prep_function(train_fold)
      validation_fold <- prep_function(validation_fold)
    }
    
    # Train model on fold
    fold_model <- fit(workflow, data = train_fold)
    
    # Generate predictions on validation fold
    preds <- predict(fold_model, new_data = validation_fold, type = "prob")$.pred_1
    
    tibble(
      obs = as.numeric(as.character(validation_fold$EVT)),
      pred = preds,
      fold = i
    )
  })
}

# Prep data type function for CART and RF

prep_for_trees <- function(data) {
  data %>%
    mutate(
      vigilance_4ISS = factor(vigilance_4ISS),
      gaze_head_deviation_4ISS = factor(gaze_head_deviation_4ISS),
      hemiparesis_4ISS = factor(hemiparesis_4ISS),
      aphasia_dysarthria_4ISS = factor(aphasia_dysarthria_4ISS),
      gender = factor(gender),
      medication_anticoags = factor(medication_anticoags),
      exclusion_TOO = factor(exclusion_TOO),
      pupil_reaction_abnormal = factor(pupil_reaction_abnormal),
      EVT = factor(EVT, levels = c("1", "0"))
    )
}

```

**Generate OOF-Predictions for All Models**

```{r}

# other models do not require special data preparation
oof_knn <- get_oof_predictions(final_knn_workflow, cv_folds, train_data)
oof_xgb <- get_oof_predictions(final_xgb_workflow, cv_folds, train_data)
oof_nn  <- get_oof_predictions(final_nn_workflow, cv_folds, train_data)

# CART and RF require factor variables
oof_cart <- get_oof_predictions(final_cart_workflow, cv_folds, train_data, 
                                 prep_function = prep_for_trees)
oof_rf   <- get_oof_predictions(final_rf_workflow, cv_folds, train_data, 
                                 prep_function = prep_for_trees)

# 4I-SS with min-max normalization
oof_4iss <- map_dfr(1:nrow(cv_folds), function(i) {
  
  split <- cv_folds$splits[[i]]
  train_fold <- training(split)
  validation_fold <- testing(split)
  
# Calculate 4I-SS 
 score_train        <- train_fold$hemiparesis_4ISS +
                       train_fold$vigilance_4ISS +
                       train_fold$gaze_head_deviation_4ISS +
                       train_fold$aphasia_dysarthria_4ISS
 
 score_validation   <- validation_fold$hemiparesis_4ISS +              
                       validation_fold$vigilance_4ISS +
                       validation_fold$gaze_head_deviation_4ISS + 
                       validation_fold$aphasia_dysarthria_4ISS
   
  # Min-max normalization using just train fold parameters
  min_s <- min(score_train)
  max_s <- max(score_train)
  
  tibble(
    obs = as.numeric(as.character(validation_fold$EVT)),
    pred = (score_validation - min_s) / (max_s - min_s),
    fold = i
  )
})

```

```{r}

# Fit platt calibration models on OOF-Predictions

calib_knn_scaled  <- glm(obs ~ pred,
                        data = oof_knn,
                        family = binomial())

calib_cart_scaled <- glm(obs ~ pred,
                        data = oof_cart,
                        family = binomial())

calib_xgb_scaled  <- glm(obs ~ pred,
                        data = oof_xgb,
                        family = binomial())

calib_rf_scaled   <- glm(obs ~ pred,
                        data = oof_rf,
                        family = binomial())

calib_nn_scaled   <- glm(obs ~ pred,
                        data = oof_nn,
                        family = binomial())

```

```{r}

# Predict platt calibrated probabilities 

calibration_data_test$KNN_cal_scaled <- predict(
  calib_knn_scaled,
  newdata = tibble(pred = calibration_data_test$KNN),
  type = "response"
)

calibration_data_test$CART_cal_scaled <- predict(
  calib_cart_scaled,
  newdata = tibble(pred = calibration_data_test$CART),
  type = "response"
)

calibration_data_test$XGB_cal_scaled <- predict(
  calib_xgb_scaled,
  newdata = tibble(pred = calibration_data_test$XGB),
  type = "response"
)

calibration_data_test$RF_cal_scaled <- predict(
  calib_rf_scaled,
  newdata = tibble(pred = calibration_data_test$RF),
  type = "response"
)

calibration_data_test$NN_cal_scaled <- predict(
  calib_nn_scaled,
  newdata = tibble(pred = calibration_data_test$NN),
  type = "response"
)

```

**Calibration of Stroke Scales**

Standard stroke scales, such as the 4I-SS, typically output ordinal point values (e.g., 0–7 points) rather than explicit probabilities. Decision Curve Analysis (DCA) fundamentally relies on **threshold probabilities** to calculate the Net Benefit. Consequently, raw score points cannot be directly plotted on the decision curve as they are incompatible with the probability scale of the x-axis.

Furthermore, clinical scores are often **non-linear** in their risk prediction. A score of 5 on a 10-point scale does indicate a 50% probability. For instance, the risk may remain negligible for scores 0–3 (low risk) and rise sharply at a specific "tipping point" (e.g., score $\ge$ 4), what would be a sigmoidal rather than linear relationship.

L**ogistic recalibration** can and should also be applied to clinical scores.

```{r}

# Fit logistic calibration models on train data
calib_4iss_scaled <- glm(obs ~ pred,
                        data = oof_4iss,
                        family = binomial())


# Apply calibration to test data
calibration_data_test$score_4ISS_cal_scaled <- predict(
  calib_4iss_scaled,
  newdata = tibble(pred = calibration_data_test$score_4ISS),
  type = "response"
)

```

**Export fitted calibration models for Brier score calculation in ML scripts**

```{r}

saveRDS(calib_knn_scaled,  here("Code",
                                "Machine Learning",
                                "models",
                                "calib_knn_scaled.rds"))

saveRDS(calib_cart_scaled, here("Code",
                                "Machine Learning",
                                "models",
                                "calib_cart_scaled.rds"))

saveRDS(calib_xgb_scaled,  here("Code",
                                "Machine Learning",
                                "models",
                                "calib_xgb_scaled.rds"))

saveRDS(calib_rf_scaled,   here("Code",
                                "Machine Learning",
                                "models",
                                "calib_rf_scaled.rds"))

saveRDS(calib_nn_scaled,   here("Code",
                                "Machine Learning",
                                "models",
                                "calib_nn_scaled.rds"))

saveRDS(calib_4iss_scaled, here("Code",
                                "Machine Learning",
                                "models",
                                "calib_4iss_scaled.rds"))

```

## Calibration Curve Before Scaling

**Standard configurations**

```{r}

# Uncalibrated calibration plot data
models <- c("KNN", "CART", "XGB", "RF", "NN", "4I-SS")

calibration_cols_unscaled <- c(
  "KNN"   = "KNN",
  "CART"  = "CART",
  "XGB"   = "XGB",
  "RF"    = "RF",
  "NN"    = "NN",
  "4I-SS" = "score_4ISS"
)

# Define model colors
model_colors <- c(
  "KNN"    = "#0072B5E6",
  "CART"   = "#E18727E6",
  "XGB"    = "#20854EE6",
  "RF"     = "#7876B1E6",
  "NN"     = "#6F99ADE6",
  "4I-SS"  = "#FFDC91E6"
)

```

**Density plot of predicted and unscaled probabilities**

The dataset is highly imbalanced, which directly affects the distribution of predicted probabilities. This imbalance must be considered when interpreting and processing calibration curves and selecting scaling methods.

```{r}

probabilities_long_format_uncal <- map_dfr(models, function(m) {
  tibble(
    model = m,
    probabilities = calibration_data_test[[ calibration_cols_unscaled[[m]] ]])}) %>%
    filter(!is.na(probabilities)) %>% 
    mutate(model = factor(model, levels = models))

```

```{r}

density_plot_probabilities_uncalibrated <- ggplot(
  probabilities_long_format_uncal,
  aes(x = probabilities, color = model)) +
  geom_density(linewidth = 0.9, adjust = 1.0) +
  scale_color_manual(values = model_colors, name = "Model") +
  scale_x_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    labels = scales::percent
  ) +
  labs(x = "Predicted Probability", y = "Density") +
  theme_minimal() +
  theme(
    legend.position = "right",
    axis.title.x = element_text(face = "bold",
                                size = 10,
                                margin = margin(t = 10)),
    axis.title.y = element_text(face = "bold",
                                size = 10,
                                margin = margin(r = 10)),
    panel.grid.minor = element_blank()
  )

print(density_plot_probabilities_uncalibrated)

```

**Calibration curve before scaling**

```{r}
#| message: false

# Calibration curve function (equal number of cases per bin)
calc_calibration_quantile_bins <- function(pred, obs, n_bins = 5) {
  df <- data.frame(pred = pred, obs = obs)
  df <- df[!is.na(df$pred) & !is.na(df$obs), ]

  # Assign each case to a quantile bin
  df$bin_id <- ntile(df$pred, n_bins)

  out <- df %>%
    group_by(bin_id) %>%
    summarise(
      mean_pred = mean(pred),
      mean_obs  = mean(obs),
      n = n(),            
      .groups = "drop"
    ) %>%
    arrange(bin_id)

  # Create a readable bin label
  out$bin <- factor(paste0("Q", out$bin_id), levels = paste0("Q", 1:n_bins))

  out %>%
    select(bin, mean_pred, mean_obs, n) 
}

# Build calibration summary (uncalibrated)
cal_results_before <- map_dfr(models, function(m) {
  tmp <- calc_calibration_quantile_bins(
    pred = calibration_data_test[[ calibration_cols_unscaled[[m]] ]],
    obs  = calibration_data_test$obs,
    n_bins = 5
  )
  tmp$model <- m
  tmp
}) %>%
  mutate(model = factor(model, levels = models)) %>%
  arrange(model, mean_pred)

```

```{r}

# Plot (uncalibrated)
calibration_plot_before <- ggplot(
  cal_results_before,
  aes(x = mean_pred, y = mean_obs, color = model, group = model)
) +
  geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    color = "grey40",
    linewidth = 0.8
  ) +
  geom_line(linewidth = 0.9) +
  geom_point(alpha = 0.8) +
  scale_color_manual(values = model_colors, name = "Model") +
  scale_x_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    labels = scales::percent
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    labels = scales::percent
  ) +
  labs(
    x = "Mean Predicted Probability",
    y = "Observed Event Rate",
    title = NULL
  ) +
  coord_fixed(ratio = 1) +
  theme_minimal() +
  theme(
    legend.position = "right",
    
    # Legend text styling
    legend.title = element_text(face = "bold", size = 10),
    legend.text = element_text(size = 9),
    
    axis.title.x = element_text(face = "bold",
                                size = 10,
                                margin = margin(t = 10)),
    axis.title.y = element_text(face = "bold",
                                size = 10,
                                margin = margin(r = 10)),
    panel.grid.minor = element_blank()
  )

print(calibration_plot_before)

```

**Safe the calibation plot before scaling as png.**

```{r}

ggsave(
  filename = "~/Downloads/calibration_plot_before.png",
  plot = calibration_plot_before,       
  device = "png",
  width = 8,            
  height = 5,           
  units = "in",
  dpi = 600,              
  bg = "white"        
)

```

## Calibration Curve After Scaling

**Density plot of predicted and unscaled probabilities**

The dataset is highly imbalanced, which directly affects the distribution of predicted probabilities. This imbalance must be considered when interpreting and processing calibration curves and selecting scaling methods.

**Standard configurations**

```{r}

calibration_cols_scaled <- c(
  "KNN"   = "KNN_cal_scaled",
  "CART"  = "CART_cal_scaled",
  "XGB"   = "XGB_cal_scaled",
  "RF"    = "RF_cal_scaled",
  "NN"    = "NN_cal_scaled",
  "4I-SS" = "score_4ISS_cal_scaled"
)

```

```{r}

probabilities_long_format_cal <- map_dfr(models, function(m) {
  tibble(
    model = m,
    probabilities = calibration_data_test[[ calibration_cols_scaled[[m]] ]]
  )
}) %>%
  filter(!is.na(probabilities)) %>%
  mutate(model = factor(model, levels = models))

```

```{r}

density_plot_probabilities_calibrated <- ggplot(
  probabilities_long_format_cal,
  aes(x = probabilities, color = model)
) +
  geom_density(linewidth = 0.9, adjust = 1.0) +
  scale_color_manual(values = model_colors, name = "Model") +
  scale_x_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    labels = scales::percent
  ) +
  labs(x = "Predicted Probability (Calibrated)", y = "Density") +
  theme_minimal() +
  theme(
    legend.position = "right",
    axis.title.x = element_text(face = "bold", size = 10, margin = margin(t = 10)),
    axis.title.y = element_text(face = "bold", size = 10, margin = margin(r = 10)),
    panel.grid.minor = element_blank()
  )

print(density_plot_probabilities_calibrated)

```

**Calibration curve after platt scaling**

```{r}
#| warning: false

# Function to calculate calibration curve data (equal number of cases per bin)
calc_calibration_after <- function(pred, obs, n_bins = 5) {

  df <- tibble(pred = pred, obs = obs) %>%
    filter(!is.na(pred), !is.na(obs))

  # Assign each case to a quantile bin (1..n_bins)
  df <- df %>%
    mutate(bin_id = ntile(pred, n_bins)) %>%
    group_by(bin_id) %>%
    summarise(
      mean_pred = mean(pred, na.rm = TRUE),
      mean_obs  = mean(obs,  na.rm = TRUE),
      n = n(),                    
      .groups = "drop"
    ) %>%
    arrange(bin_id)

  # Create readable bin labels
  df$bin <- factor(paste0("Q", df$bin_id), levels = paste0("Q", 1:n_bins))

  df %>%
    select(bin, mean_pred, mean_obs, n) 
}


cal_results_after <- map_dfr(models, function(m) {
  tmp <- calc_calibration_after(
    pred = calibration_data_test[[ calibration_cols_scaled[[m]] ]],
    obs  = calibration_data_test$obs,
    n_bins = 5
  )
  tmp$model <- m
  tmp
})

# Set factor levels to control legend order
cal_results_after <- cal_results_after %>%
  mutate(model = factor(model, levels = models)) %>%
  arrange(model, mean_pred)

```

```{r}

# Plot calibration curves 
calibration_plot_after <- ggplot(
  cal_results_after,
  aes(x = mean_pred, y = mean_obs, color = model, group = model)
) +
  geom_abline(intercept = 0,
              slope = 1,
              linetype = "dashed",
              color = "grey40",
              linewidth = 0.8) +
  geom_line(linewidth = 0.9) +
  geom_point(alpha = 0.8) +
  scale_color_manual(values = model_colors, name = "Model") +
  scale_x_continuous(limits = c(0, 1),
                     breaks = seq(0, 1, 0.2),
                     labels = scales::percent) +
  scale_y_continuous(limits = c(0, 1),
                     breaks = seq(0, 1, 0.2),
                     labels = scales::percent) +
  labs(
    x = "Mean Predicted Probability",
    y = "Observed Event Rate",
    title = NULL
  ) +
  coord_fixed(ratio = 1) +
  theme_minimal() +
  theme(
    legend.position = "right",
    
    # Legend text styling
    legend.title = element_text(face = "bold", size = 10),
    legend.text = element_text(size = 9),
    
    axis.title.x = element_text(face = "bold",
                                size = 10,
                                margin = margin(t = 10)),
    axis.title.y = element_text(face = "bold",
                                size = 10,
                                margin = margin(r = 10)),
    
    panel.grid.minor = element_blank()
  )

print(calibration_plot_after)

```

**Safe the calibration plot after scaling as png.**

```{r}

ggsave(
  filename = "~/Downloads/calibration_plot_after.png",
  plot = calibration_plot_after,       
  device = "png",
  width = 8,            
  height = 5,           
  units = "in",
  dpi = 600,              
  bg = "white"        
)

```

## Combined Presentation

Use `cowplot` for combination

```{r}

# Extract legend from one of the plots
legend_grob <- cowplot::get_legend(
  calibration_plot_before +
    theme(
      legend.position = "top",
      legend.title = element_text(face = "bold", size = 10),
      legend.text  = element_text(size = 9)
    )
)

# Remove legends from both plots
p1 <- calibration_plot_before +
  labs(title = "Before Platt Scaling") +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 11, hjust = 0.5)
  )

p2 <- calibration_plot_after +
  labs(title = "After Platt Scaling") +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 11, hjust = 0.5)
  )

# Combine: legend on top, plots below
calibration_final <- wrap_elements(legend_grob) / (p1 | p2) +
  plot_layout(heights = c(0.08, 1))

print(calibration_final)

```

**Safe the combined calibration plot as png.**

```{r}

ggsave(
  filename = "~/Downloads/calibration_combined.png",
  plot = calibration_final,
  device = "png",
  width = 9.5,
  height = 5,
  units = "in",
  dpi = 600,
  bg = "white"
)

```

## Quantitative Evaluation

Compare quantitative calibration metrics in a table.

```{r}
#| warning: false

calc_calib_metrics <- function(data, prob_col, obs_col = "obs", n_bins = 5) {
  
  if (!prob_col %in% colnames(data)) return(NULL)
  
  probs <- data[[prob_col]]
  obs   <- data[[obs_col]]
  
  brier <- mean((probs - obs)^2)
  
  logit_probs <- qlogis(pmin(pmax(probs, 1e-7), 1 - 1e-7))
  calib_model <- tryCatch(glm(obs ~ logit_probs, family = binomial), error = function(e) NULL)
  
  if (!is.null(calib_model)) {
    intercept <- coef(calib_model)[1]
    slope     <- coef(calib_model)[2]
  } else {
    intercept <- NA
    slope <- NA
  }
  
  bins <- cut(probs, breaks = seq(0, 1, length.out = n_bins + 1), include.lowest = TRUE)
  
  bin_stats <- tibble(prob = probs, obs = obs, bin = bins) %>%
    group_by(bin) %>%
    summarise(mean_pred = mean(prob), mean_obs = mean(obs), n = n(), .groups = "drop") %>%
    mutate(abs_diff = abs(mean_pred - mean_obs), weight = n / length(probs))
  
  ece <- sum(bin_stats$weight * bin_stats$abs_diff, na.rm = TRUE)
  mce <- max(bin_stats$abs_diff, na.rm = TRUE)
  
  ici <- tryCatch({
    loess_fit <- loess(obs ~ probs, degree = 1, span = 0.75)
    mean(abs(fitted(loess_fit) - probs), na.rm = TRUE)
  }, error = function(e) NA)
  
  tibble(
    "Brier Score" = brier,
    "Calib. Intercept" = intercept,
    "Calib. Slope" = slope,
    "ECE" = ece,
    "MCE" = mce,
    "ICI" = ici
  )
}

```

**Unscaled vs. Scaled**

```{r}
#| warning: false

model_map <- list(
  "KNN"   = c(raw = "KNN",        platt = "KNN_cal_scaled"),
  "CART"  = c(raw = "CART",       platt = "CART_cal_scaled"),
  "XGB"   = c(raw = "XGB",        platt = "XGB_cal_scaled"),
  "RF"    = c(raw = "RF",         platt = "RF_cal_scaled"),
  "NN"    = c(raw = "NN",         platt = "NN_cal_scaled"),
  "4I-SS" = c(raw = "score_4ISS", platt = "score_4ISS_cal_scaled")
)

results_list <- list()

for (model_name in names(model_map)) {
  cols <- model_map[[model_name]]
  for (type in names(cols)) {
    metrics <- calc_calib_metrics(calibration_data_test, cols[type])
    if (!is.null(metrics)) {
      results_list[[paste0(model_name, "_", type)]] <- metrics %>%
        mutate(
          Model = model_name,
          Method = ifelse(type == "raw", "Unscaled", "Scaled")
        )
    }
  }
}

comparison_table <- bind_rows(results_list) %>%
  select(Model, Method, `Brier Score`, `Calib. Slope`, `Calib. Intercept`, ECE, MCE, ICI)

print(comparison_table)

```

# Decision Curve Analysis

## Background

**This Paper Explains DCA´s**

Vickers, A. J., van Calster, B., & Steyerberg, E. W. (2019). A simple, step-by-step guide to interpreting decision curve analysis. *Diagnostic and prognostic research*, *3*(1), 18.

https://doi.org/10.1186/s41512-019-0064-7

**DCA is not easy to understand**

*"a well-respected epidemiologist claimed that he had yet to find more than a couple of people in the world who could explain what decision curves meant and that he himself was not clear on their interpretation. We have also attended meetings where presenters have shown a decision curve slide and then commented that they themselves did not actually understand it."*

**What is Net-Benefit?**

$$
\text{Net Benefit} = \frac{\text{TP} - \text{FP} \times w}{N}
$$

-   **TP** = True positives

-   **FP** = False positives

-   **w** = harm weight of false positive diagnoses (relative to benefit of true positives)

-   N = total number of patients

**What is the Net Benefit?**

"The “net” in “net benefit” is the same as in “net profit,” that is, income minus expenditure. If, say, a wine importer buys €1m of wine from France and sells it in the USA for \$1.5m, then if the exchange rate is €1 to \$1.25, the net profit is income in dollars (1.5m) − expenditure in euros (1m) × exchange rate (1.25) = \$250,000. Leaving aside, for the sake of simplicity, the issue of risk and the time and trouble to trade, this is equivalent to being given \$250,000 without having to do any trading" (Vickers et. all)

The unit of net benefit is true positives. A net benefit of 0.07, for instance, means “7 true positives for every 100 patients in the target population.

**What does the EVT Probability (here also weight (w)) mean?**

The risk of 10% is an odds of 1:9, so in using a threshold probability of 10% and tells us that undertriage is 9 times worse than overtriage. This can be interpreted as the “number-needed-to-test,” that is, 10% is a number-needed-to-test of 10. In other words, to find one true positive, we accept 9 false positives.

**Intervention for all line:**

The curve intersects both the vertical y-axis and the horizontal x-axis at the prevalence.

-   **Y-axis (threshold = 0):** At this point, the harm of a false positive is weighted as zero, meaning it carries no penalty. When we treat everyone without testing, only those who actually have the disease (equal to the prevalence) benefit. The false positives cause no weighted harm and therefore don't reduce the net benefit, leaving it equal to the prevalence.

-   **X-axis (Net benefit = 0):** The "Treat all" curve intersects the x-axis where the net benefit equals zero. At a prevalence of 10%, this occurs at threshold pt = 0.10. Here, we gain benefit from treating 10 truly diseased patients per 100, but we also incur harm from 90 false positives per 100. At this threshold, one false positive is weighted 1/9th as heavily as one true positive (pt/(1-pt) = 0.10/0.90). Therefore, the weighted harm from 90 false positives (90 × 1/9 = 10 units) exactly cancels out the benefit from 10 true positives (10 units), resulting in a net benefit of zero.

**Intervention for none line:**

The curve is horizontal at a net benefit of 0 because no patients receive treatment. Since there are neither true positives nor false positives, the net benefit is 0 across all threshold probabilities.

**Why is “intervention for all” or “intervention for none” a relevant comparison?**

"Intervening for all or no patients, irrespective of test or model results are reasonable clinical strategies in many scenarios. A test or model must be found superior to both of these strategies to justify being used in clinical practice" (Vickers et. all)

**Influence of Sensitivity and Specificity on the DCA:**

Although both sensitivity and specificity influence the net benefit, their relative importance shifts across the threshold spectrum: at lower thresholds (left side of the curve), high sensitivity is more important, whereas at higher thresholds (right side), high specificity becomes more critical—because as the threshold increases, false positives carry greater weight, and specificity directly reduces false positives.

**Recommended by Journals**

DCA has been recommended by top journals, including BMJ, JAMA and PLoS Medicine

## Calculate Net Benefit

Calculate the net benefit for a threshold range of 1% to 99%

```{r}
#| message: false

dca_results <- dcurves::dca(
  data = calibration_data_test,
  formula = obs ~ KNN_cal_scaled + CART_cal_scaled + XGB_cal_scaled + RF_cal_scaled + NN_cal_scaled + score_4ISS_cal_scaled,
  thresholds = seq(0.01, 0.99, by = 0.01), 
  label = list(
    KNN_cal_scaled  = "KNN",
    CART_cal_scaled = "CART",
    XGB_cal_scaled  = "XGB",
    RF_cal_scaled   = "Random Forest",
    NN_cal_scaled   = "Neural Network",
    score_4ISS_cal_scaled = "4I-SS"
  )
)

```

## Plot DCA

```{r}

# Preparation
plot_data <- as_tibble(dca_results) %>%
  mutate(
    # Rename default dca labels 
    label = dplyr::case_match(label,
      "Treat All"  ~ "Intervention for all",
      "Treat None" ~ "Intervention for none",
      .default = label
    ),
    # Set the exact order for the legend
    label = factor(label, levels = c(
      "KNN",
      "CART",
      "XGB",
      "Random Forest",
      "Neural Network",
      "4I-SS",
      "Intervention for all",
      "Intervention for none"
    )),
    # Convert Net Benefit to "per 100 patients"
    net_benefit = net_benefit * 100 
  )

# Separate data to control layers individually
baselines <- c("Intervention for all", "Intervention for none")
data_models <- plot_data %>% filter(!label %in% baselines)
data_baselines <- plot_data %>% filter(label %in% baselines)

# DCA Plotting
dca_plot <- ggplot() +
  
  # Layer 1: ML Models
  geom_line(
    data = data_models,
    aes(x = threshold, y = net_benefit, color = label),
    linewidth = 0.7,
    linetype = "solid"
  ) +
  
  # Layer 2: Reference Strategies
  geom_line(
    data = data_baselines,
    aes(x = threshold, y = net_benefit, linetype = label),
    linewidth = 0.7,
    color = "grey40" 
  ) +
  
  # Colors for the models
  scale_color_manual(
    name = "Preclinical Triage Strategies",
    values = c(
      "KNN"            = "#0072B5E6",
      "CART"           = "#E18727E6",
      "XGB"            = "#20854EE6",
      "Random Forest"  = "#7876B1E6",
      "Neural Network" = "#6F99ADE6",
      "4I-SS"          = "#FFDC91E6"
    )
  ) +
  
  # Linetypes
  scale_linetype_manual(
    name = "Reference Strategies",
    values = c(
      "Intervention for all"  = "dashed",
      "Intervention for none" = "solid"
    )
  ) +
  
  # Define Legend Order and Layout
  guides(
    color = guide_legend(order = 1),    # Models lines on top
    linetype = guide_legend(order = 2)  # References lines behind
  ) +
  
  # Axis Scaling
  coord_cartesian(ylim = c(-1.5, 8), xlim = c(0, 0.50)) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  
  # Labels
  labs(
    x = "Minimize Undertriage      <-       Probability of EVT       ->      Minimize Overtriage",
    y = "Net Benefit (per 100 patients)",
    title = NULL
  ) +
  
  # Theme Settings
  theme_minimal() +
  theme(
    # Legend positioning
    legend.position = "right", 
    legend.box = "vertical", 
    legend.box.just = "left",
    legend.margin = margin(t = 0, b = 0),
    legend.spacing.y = unit(0.5, "cm"),
    
    # Legend text styling
    legend.title = element_text(face = "bold",
                                size = 10),
    legend.text = element_text(size = 9),
    
    # Axis titles 
    axis.title.x = element_text(face = "bold",
                                size = 10,
                                margin = margin(t = 10)),
    axis.title.y = element_text(face = "bold",
                                size = 10,
                                margin = margin(r = 8))
  )

# Print Plot
print(dca_plot)

```

**Safe the DCA as png.**

```{r}

ggsave(
  filename = "~/Downloads/DCA_analysis_all_models.png",
  plot = dca_plot,       
  device = "png",
  width = 8,            
  height = 5,           
  units = "in",
  dpi = 600,              
  bg = "white"        
)

```

# Bootstrapped Performance Distributions

This section compares the bootstrap distributions and confidence intervals of performance metrics across all models using ridgeline plots. We use 5,000 bootstrap resamples with BCa (bias-corrected and accelerated) confidence intervals, consistent with the methodology applied in individual model scripts.

## Data Preparation

We define a custom function to calculate all performance metrics for each bootstrapped dataset. This function will be called by \`boot::boot()\` for each of the 5,000 resamples.

```{r}

# Define model colors matching the ROC curves
colors_nejm_alpha90 <- c(
       "#BC3C29E6",
       "#E18727E6",
       "#0072B5E6",
       "#20854EE6", 
       "#7876B1E6",
       "#6F99ADE6",
       "#FFDC91E6",
       "#EE4C97E6"
       )

```

**Choose Model Thresholds from the ML-Scripts**

```{r}

model_thresholds <- c(
  "OneR"  = 0.5,
  "KNN"   = 0.5,
  "CART"  = 0.4,  
  "XGB"   = 0.4,   
  "RF"    = 0.4,   
  "NN"    = 0.45,   
  "4I-SS" = 2/7   # 2/7 equals Score > 2
)

```

Map raw-model column names -\> calibrated column names (only for Brier)

```{r}

cal_map <- c(
  "KNN"   = "KNN_cal_scaled",
  "CART"  = "CART_cal_scaled",
  "XGB"   = "XGB_cal_scaled",
  "RF"    = "RF_cal_scaled",
  "NN"    = "NN_cal_scaled",
  "4I-SS" = "score_4ISS_cal_scaled"
)

```

**Create a Single Master Table with all Information for the Bootstrapping**

We create a dataframe with the label (ground truth), all observations of the test data, and the predicted probability for each model.

```{r}

# Combines truth and probabilities/scores from all models
data_for_bootstrapping <- as.data.frame(
  list(
    EVT     = label_test,
    OneR    = model_test_scores$OneR,
    CART    = model_test_scores$CART,
    KNN     = model_test_scores$KNN,
    XGB     = model_test_scores$XGB,
    RF      = model_test_scores$RF,
    NN      = model_test_scores$NN,
    `4I-SS` = model_test_scores$`4I-SS`,
    
# calibrated probs (only for Brier)

    CART_cal_scaled       = calibration_data_test$CART_cal_scaled,
    KNN_cal_scaled        = calibration_data_test$KNN_cal_scaled,
    XGB_cal_scaled        = calibration_data_test$XGB_cal_scaled,
    RF_cal_scaled         = calibration_data_test$RF_cal_scaled,
    NN_cal_scaled         = calibration_data_test$NN_cal_scaled,
    score_4ISS_cal_scaled = calibration_data_test$score_4ISS_cal_scaled
  ),
  check.names = FALSE
)


# Check the data structure
head(data_for_bootstrapping)

```

## Define Custom Function

We define a custom function to calculate all performance metrics for each bootstrapped dataset. This function will be called by \`boot::boot()\` for each of the 5,000 resamples.

```{r}

calc_boot_metrics_all_models <- function(data, indices) {

  # Bootstrap resample
  d <- data[indices, ]

  # True labels (factor)
  truth_raw <- factor(as.character(d$EVT), levels = c("0", "1"))

  # True labels (numeric)
  truth_num <- ifelse(truth_raw == "1", 1, 0)

  # Store metrics
  results <- c()

  for (model in names(model_thresholds)) {

      # Model scores
      probs_raw <- d[[model]]

      # Decision threshold
      thresh <- model_thresholds[[model]]

      # Class predictions
      pred_class <- factor(ifelse(probs_raw > thresh, "1", "0"),
                           levels = c("0", "1"))

      # Create Confusion matrix (calucate most metrics)
      cm <- caret::confusionMatrix(pred_class, truth_raw, positive = "1")

      # PR input
      pr_df <- tibble(truth = truth_raw, prob = probs_raw)

      # ROC AUC
      aucroc <- as.numeric(pROC::auc(pROC::roc(
        response  = truth_raw,
        predictor = probs_raw,
        levels    = c("0", "1"),
        direction = "<",
        quiet     = TRUE
      )))

      # PR curve
      pr_obj <- PRROC::pr.curve(
        scores.class0 = probs_raw[truth_raw == "1"],
        scores.class1 = probs_raw[truth_raw == "0"]
      )

      # PR AUC
      AUPRC <- as.numeric(pr_obj$auc.integral)

      # Brier score
      if (model == "OneR") {
        brier <- NA_real_
      } else {
        cal_col <- unname(cal_map[model])
        if (is.na(cal_col) || !(cal_col %in% names(d))) {
          brier <- NA_real_
        } else {
          cal_probs <- d[[cal_col]]
          brier <- mean((truth_num - cal_probs)^2, na.rm = TRUE)
        }
      }

      # Collect outputs
      results <- c(
        results,
        AUCROC        = aucroc,
        AUPRC         = AUPRC,
        Accuracy      = as.numeric(cm$overall["Accuracy"]),
        Sensitivity   = as.numeric(cm$byClass["Sensitivity"]),
        Specificity   = as.numeric(cm$byClass["Specificity"]),
        `Brier Score` = brier,
        PPV           = as.numeric(cm$byClass["Pos Pred Value"]),
        NPV           = as.numeric(cm$byClass["Neg Pred Value"])
      )
    }

    # Return metrics
    results
  }

```

We perform the bootstrap simulation with 5,000 replicates.

```{r}

set.seed(456)

# Execute bootstrapping with 5,000 replicates using the statistic function
boot_results_all <- boot::boot(
  data = data_for_bootstrapping,
  statistic = calc_boot_metrics_all_models,
  R = 5000
)

```

Prepare bootstrap samples for plotting

```{r}

# Define metrics names
calc_order_names <- c("AUCROC",
                      "AUPRC",
                      "Accuracy",
                      "Sensitivity",
                      "Specificity",
                      "Brier Score",
                      "PPV",
                      "NPV"
)

model_names <- names(model_thresholds)

# Create column mapping
col_mapping <- expand.grid(Metric = calc_order_names, Model = model_names)

# Convert matrix to a dataframe
boot_samples_df <- as.data.frame(boot_results_all$t)

# Define column names
colnames(boot_samples_df) <- paste(col_mapping$Model,
                                   col_mapping$Metric,
                                   sep = "_")

# Define visualization order
desired_plot_order <- c("AUCROC",
                        "AUPRC",
                        "Accuracy",
                        "Sensitivity",
                        "Specificity",
                        "PPV",
                        "NPV",
                        "Brier Score")

# Transform to long format and set factor levels for plotting
boot_results_all_long <- boot_samples_df %>%
  pivot_longer(everything(),
               names_to = "Model_Metric",
               values_to = "Value") %>%
  separate(Model_Metric,
           into = c("Model", "Metric"),
           sep = "_",
           extra = "merge") %>%
  mutate(
    Model  = factor(Model, levels = model_names),
    Metric = factor(Metric, levels = desired_plot_order)
  )

```

**Calculate BCa´s Confidence Intervals**

```{r}

# BCa CI function that returns lower and upper bounds
get_bca_limits <- function(boot_obj, index, conf) {

  t_col <- boot_obj$t[, index]
  t_ok  <- t_col[is.finite(t_col)]
  if (length(t_ok) < 10) return(c(NA_real_, NA_real_))

  if (!all(is.finite(t_col))) {
    probs <- c((1-conf)/2, 1-(1-conf)/2)
    return(as.numeric(stats::quantile(t_ok, probs = probs, na.rm = TRUE)))
  }

  ci <- tryCatch(
    boot::boot.ci(boot_obj, type = "bca", conf = conf, index = index),
    error = function(e) NULL
  )

  if (is.null(ci) || is.null(ci$bca) || anyNA(ci$bca[4:5])) {
    probs <- c((1-conf)/2, 1-(1-conf)/2)
    return(as.numeric(stats::quantile(t_ok, probs = probs, na.rm = TRUE)))
  }

  ci$bca[4:5] # lower, upper
}

get_bca_and_mean <- function(boot_obj, index) {

  # Calculate the bootstrap mean to include it in the plot
  t_col <- boot_obj$t[, index]
  t_ok  <- t_col[is.finite(t_col)]
  boot_mean <- mean(t_ok, na.rm = TRUE)

  # Calculate uncertainty bands for BCa 95/80/50
  ci95 <- get_bca_limits(boot_obj, index, 0.95)
  ci80 <- get_bca_limits(boot_obj, index, 0.80)
  ci50 <- get_bca_limits(boot_obj, index, 0.50)

  list(
    mean = boot_mean * 100,
    q025 = ci95[1] * 100, q975 = ci95[2] * 100,
    q10  = ci80[1] * 100, q90  = ci80[2] * 100,
    q25  = ci50[1] * 100, q75  = ci50[2] * 100
  )
}

# Remove Brier Score Values for OneR and the 4I-SS
# OneR and the 4I-SS are scales and no probabilities, so the Brier Score is not defined
excluded_brier_models <- c("OneR")

# Build intervals per Model and Metric
interval_data <- purrr::map_dfr(seq_len(nrow(col_mapping)), function(i) {

  # Keep the row so the model still appears in the Brier facet,
  # but do not compute BCa; show NA in the Mean column.
  if (col_mapping$Metric[i] == "Brier Score" &&
      col_mapping$Model[i] %in% excluded_brier_models) {

    return(tibble::tibble(
      Model  = col_mapping$Model[i],
      Metric = col_mapping$Metric[i],
      mean   = NA_real_,
      q025   = NA_real_, q975 = NA_real_,
      q10    = NA_real_, q90  = NA_real_,
      q25    = NA_real_, q75  = NA_real_
    ))
  }

  res <- get_bca_and_mean(boot_results_all, index = i)

  tibble::tibble(
    Model  = col_mapping$Model[i],
    Metric = col_mapping$Metric[i],
    mean   = res$mean,
    q025   = res$q025, q975 = res$q975,
    q10    = res$q10,  q90  = res$q90,
    q25    = res$q25,  q75  = res$q75
  )
}) %>%
  dplyr::mutate(
    Model  = factor(Model, levels = rev(model_names)),
    Metric = factor(Metric, levels = desired_plot_order),
    mean_label = dplyr::case_when(
      is.na(mean) ~ "NA",
      Metric == "Brier Score" ~ sprintf("%.1f", mean),
      TRUE ~ sprintf("%.0f", mean)
    )
  )

# Prepare long bootstrap data (for densities + axis ranges)
plot_data <- boot_results_all_long %>%
  dplyr::mutate(
    Model  = factor(Model, levels = rev(model_names)),
    Metric = factor(Metric, levels = desired_plot_order),
    perf   = Value * 100
  )

# Remove rows for OneR only if Metric is Brier Score
plot_data <- plot_data %>%
  dplyr::filter(!(Metric == "Brier Score" & Model %in% excluded_brier_models))

# Per-metric x-range for text and headers
metric_meta <- plot_data %>%
  dplyr::group_by(Metric) %>%
  dplyr::summarise(
    min_val = min(perf, na.rm = TRUE),
    max_val = max(perf, na.rm = TRUE),
    range   = max_val - min_val,
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    range = ifelse(range == 0 | is.na(range), 1, range)  # avoid zero-range issues
  )

interval_data <- interval_data %>%
  dplyr::left_join(metric_meta, by = "Metric") %>%
  dplyr::mutate(text_x = min_val - (range * 0.18))

# Headers and smaller titels
n_models <- length(model_names)

header_data <- metric_meta %>%
  dplyr::mutate(
    y_pos  = n_models + 1.0,
    text_x = min_val - (range * 0.18),
    label  = "Mean"
  )

title_data <- metric_meta %>%
  dplyr::mutate(
    y_pos  = n_models + 2.2,
    text_x = min_val - (range * 0.18),
    label  = as.character(Metric)
  )

# Set same colors as model have in the ROC and PR curves
model_colors_base <- setNames(
  substr(colors_nejm_alpha90[seq_along(model_names)], 1, 7),
  model_names
)

# Band widths (95/80/50)
band_lwd_95 <- 2.6
band_lwd_80 <- 3.0
band_lwd_50 <- 3.4

```

**Remove Brier Score Values for OneR and the 4I-SS**

OneR and the 4I-SS are scales and no probabilities, so the Brier Score is not defined

```{r}

# Remove rows for OneR only if Metric is Brier Score
plot_data <- plot_data %>%
  filter(
    !(Metric == "Brier Score" & Model %in% c("OneR"))
  )

```

## Create `ggdist`-Plots

**Standardize Text Positions for 0-100 Scale**

We overwrite the text_x coordinates to ensure alignment across all facets on the fixed 0-100 scale.

```{r}

# Position the "Mean" column on the left (e.g., at x = 0%)
interval_data$text_x <- 0 
header_data$text_x   <- 0

# Position the Metric Titles also on the left (e.g., at x = 0%)
title_data$text_x    <- 0

```

**Do the Plot Using `ggplot` and `ggdist`**

```{r}
#| warning: false

ridget_plot <- ggplot(plot_data, aes(x = perf, y = Model)) +

  # A. Violine
  geom_violin(
    aes(fill = Model, color = Model), 
    alpha = 0.15,         
    scale = "width",      
    trim = TRUE,          
    linewidth = 0.1,      
    show.legend = FALSE
  ) +

  # B. Intervals
  geom_linerange(
    data = interval_data,
    aes(y = Model, xmin = q025, xmax = q975, color = Model),
    linewidth = band_lwd_95, alpha = 0.50, inherit.aes = FALSE, show.legend = FALSE
  ) +
  geom_linerange(
    data = interval_data,
    aes(y = Model, xmin = q10, xmax = q90, color = Model),
    linewidth = band_lwd_80, alpha = 0.75, inherit.aes = FALSE, show.legend = FALSE
  ) +
  geom_linerange(
    data = interval_data,
    aes(y = Model, xmin = q25, xmax = q75, color = Model),
    linewidth = band_lwd_50, alpha = 1.00, inherit.aes = FALSE, show.legend = FALSE
  ) +

  # C. Mean Point
  geom_point(
    data = interval_data,
    aes(x = mean, y = Model),
    inherit.aes = FALSE, color = "black", size = 1.2, show.legend = FALSE
  ) +

  # D. Numbers Column 
  geom_text(
    data = interval_data,
    aes(x = text_x, y = Model, label = mean_label),
    inherit.aes = FALSE,
    size = 3.2,
    family = "mono",
    fontface = "bold",
    color = "grey20",
    hjust = 0.5 
  ) +

  # E. Header "Mean" 
  geom_text(
    data = header_data,
    aes(x = text_x, y = y_pos, label = label),
    inherit.aes = FALSE,
    size = 3,
    fontface = "bold",
    color = "grey30",
    hjust = 0.5, 
    vjust = 0 
  ) +

  # F. Metric Title 
  geom_text(
    data = title_data,
    aes(x = text_x, y = y_pos, label = label),
    inherit.aes = FALSE,
    size = 4.5,        
    fontface = "bold",
    color = "black",
    hjust = 0.5,       
    vjust = 0
  ) +

  # Faceting
  facet_wrap(~Metric, scales = "free_x", ncol = 2) +

  # Colors & Fills
  scale_color_manual(values = model_colors_base) +
  scale_fill_manual(values = model_colors_base) + 
    
  # Scales
  scale_x_continuous(
     breaks = seq(0, 100, by = 10),
     limits = c(-5, 100),
     expand = expansion(mult = c(0, 0.02)) 
  ) +
    
  scale_y_discrete(
    expand = expansion(mult = c(0.05, 0.45)) 
  ) +

  coord_cartesian(xlim = c(-5, 100), clip = "off") +

  labs(x = "Bootstrapped Model Performance (%)", y = NULL) +

  theme_minimal(base_family = "sans") +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(linewidth = 0.3, color = "grey92"), 
    panel.grid.major.x = element_line(linewidth = 0.3, color = "grey92"), 
    strip.text = element_blank(), 
    strip.background = element_blank(),
    axis.text.y = element_text(face = "bold", size = 9, color = "black", hjust = 0),
    axis.text.x = element_text(size = 8, color = "grey30"),
    axis.title.x = element_text(face = "bold", size = 9, margin = margin(t = 25)), 

    panel.spacing = unit(2.5, "lines"), 
    plot.margin = margin(t = 40, r = 10, b = 10, l = 10)
  )

# Print
suppressMessages(print(ridget_plot))

```

```{r}

ggsave(
  filename = "~/Downloads/bootstrapped_distribution_comparison.png",
  plot = ridget_plot,
  width = 10,
  height = 12,
  dpi = 600,    # very high resolution      
  bg = "white"  # white backgound       
)

```
